### 2015-09-24 draft/fs_qa_v2.tex
rev-24/09/2015

\documentclass[11pt]{iopart}
\usepackage[english]{babel}
\pdfoutput=1
%\usepackage{amsmath}
\usepackage{wasysym}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{amsbsy}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{color}
\usepackage{sidecap}
\usepackage{bm}% bold math
\usepackage{tikz}
%\usepackage[backend=bibtex]{biblatex}

\usepackage[colorlinks,bookmarks=false,citecolor=blue,linkcolor=red,urlcolor=blue]{hyperref}

\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
            \node[shape=circle,draw,inner sep=2pt] (char) {#1};}}

%\usepackage{cite}

%
% WARNING !!!!
% 
% iopart.cls definition of \tableofcontents overwrites the
% short title printed every page. 
% The following redefinition of \tableofcontents fixes the
% the problem.

\catcode`@=11 % If we need private TeX macros
\renewcommand\tableofcontents{%
  \section*{\contentsname}%
  \@starttoc{toc}%
}
\catcode`@=12% '@' is no more a character
\newcommand{\bra}[1]{\langle\left.{#1}\right|}
\newcommand{\ket}[1]{\left|{#1}\right.\rangle}

\def\be{\begin{equation}}
\def\ee{\end{equation}}
\def\u{\uparrow}
\def\d{\downarrow}
\def\nm{\newmoon}
\def\fm{\fullmoon}
\def\T{\rule{0pt}{.6cm}}
\def\B{\rule[-.4cm]{0pt}{0pt}}

\begin{document}

\setlength{\parindent}{0pt}

\title{The quench action approach for integrable models: A Monte Carlo implementation}

\author{Authors}
%\address{$^1$ International School for Advanced Studies (SISSA),
%Via Bonomea 265, 34136, Trieste, Italy,
%INFN, Sezione di Trieste}

\date{\today}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

fdasfa
\end{abstract}

\maketitle

%%%%%%%%%%%%%%%%% INTRODUCTION %%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{intro}

We show that it is possible to numerically simulate the Quench Action approach 
combining Monte Carlo methods and Bethe ansatz techniques.

We focus on the situation in which the pre-quench initial state is the Neel 
state or the Majumdar-Ghosh state.

We investigate the importance of the zero-momentum strings in the Quench Action.

Without zero-momentum strings the overlap saturation rules are not valid, 
i.e., in finite size systems the vast majority of the eigenstates contain 
zero momentum strings.

The details on the eigenstates counting depend on the pre-quench initial state.

However, we show that one can restrict to the set of non-zero momentum strings. 
The fact that one neglects zero-momentum strings gives rise only to scaling 
corrections.

We also investigate the validity of the Bethe-Takahashi approximation for the 
calculation of the overlap.

In the thermodynamic limit it is natural to assume that the steady-state expectation 
values are obtained using the so-called diagonal ensemble, which is defined as 
%
\begin{equation}
\label{d-ensemble}
\langle{\mathcal O}\rangle=\sum\limits_{\alpha}|\langle\Psi_0|\alpha\rangle|^2
\langle\alpha|{\mathcal O}|\alpha\rangle. 
\end{equation}
%

It is useful to rewrite~\eref{d-ensemble} as 
%
\begin{equation}
\label{qa-prob}
\langle{\mathcal O}\rangle=\sum\limits_{\alpha}\rho^{DE}(\alpha)
\langle\alpha|{\mathcal O}|\alpha\rangle,\quad\textrm{with}\quad 
\rho^{DE}(\alpha)=\exp(-2\Re{\mathcal E}(\alpha)).
\end{equation}
%
Here $\rho^{DE}(\alpha)$ is a diagonal ensemble density matrix, and ${\mathcal E}(\alpha)
\equiv-\log\langle\alpha|\Psi_0\rangle$, with $\Re$ denoting the real part.

%%%%%%%%%%%%%%%%%%%%%%% BETHE ANSATZ FOR THE XXX CHAIN %%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Bethe ansatz solution of the Heisenberg ($XXX$) spin chain}
\label{sec:1}

Here we review some Bethe ansatz results for the spin-$\frac{1}{2}$ Heisenberg 
($XXX$) chain. Specifically, in section~\ref{sec:1.1} we introduce the model. 
Its eigenstates (Bethe states) and the Bethe equations are discussed in 
section~\ref{sec:1.2}. Section~\ref{sec:1.3} focuses on the string hypothesis 
and the so-called Bethe-Gaudin-Takahashi (BGT) equations. The form of the 
BGT equations in the thermodynamic limit is discussed in section~\ref{sec:1.4}. 
Finally, in section~\ref{sec:1.5} we provide some exact formulas for the 
local conserved charges of the model.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The spin-$\frac{1}{2}$ Heisenberg chain}
\label{sec:1.1}

The spin-$\frac{1}{2}$ isotropic Heisenberg chain ($XXX$ chain) with $L$ sites 
is defined by the Hamiltonian 
%
\begin{equation}
\label{xxx-ham}
{\mathcal H}\equiv J\sum\limits_{i=1}^L\left[\frac{1}{2}(S_i^+S^-_{i+1} 
+S_i^{-}S_{i+1}^+)+S_i^zS_{i+1}^z-\frac{1}{4}\right],  
\end{equation}
%
where $S^{\pm}_i\equiv (\sigma_i^x\pm i\sigma_i^y)/2$ are spin operators acting on the 
site $i$, $S_i^z\equiv\sigma_i^z/2$, and $\sigma^{x,y,z}_i$ the Pauli matrices. We fix 
$J=1$ and use periodic boundary conditions, identifying sites $L+1$ and $1$. The total 
magnetization $S_{T}^z\equiv\sum_iS_i^z=L/2-M$, with $M$ number of down spins (particles), 
commutes with~\eref{xxx-ham}, and it is here used to label its eigenstates.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Bethe equations and wavefunctions}
\label{sec:1.2}

In the Bethe ansatz framework~\cite{bethe-1931,taka-book} the generic eigenstate 
of~\eref{xxx-ham} (Bethe state) in the sector with $M$ particles can be written as 
%
\begin{equation}
\label{ba-eig}
|\Psi_M\rangle=\sum\limits_{1\le x_1<x_2<\dots<x_M\le L}A_M(x_1,x_2,
\dots,x_M)|x_1,x_2,\dots,x_M\rangle,
\end{equation}
%
where the sum is over the positions $\{x_i\}_{i=1}^M$ of the particles, and $A_M(x_1,
x_2,\dots,x_M)$ is the eigenstate amplitude corresponding to the particles 
being at positions $x_1,x_2,\dots, x_M$. Here $A_M(x_1,x_2,\dots, x_M)$ is 
given as 
%
\begin{equation}
\label{ba_amp}
A_M(x_1,x_2,\dots,x_M)\equiv\sum\limits_{\sigma\in S_M}\exp\Big[i
\sum\limits_{j=1}^Mk_{\sigma_j}x_j+i\sum\limits_{i<j}\theta_{\sigma_i,\sigma_j}
\Big], 
\end{equation}
%
where the outermost summation is over the permutations $S_M$ of the so-called 
quasi-momenta $\{k_\alpha\}_{\alpha=1}^M$. The two-particle scattering phases 
$\theta_{\alpha,\beta}$ are defined as 
%
\begin{equation}
\label{s_phases}
\theta_{\alpha,\beta}\equiv \frac{1}{2i}\log\Big[-\frac{e^{ik_\alpha+ik_\beta}-
2e^{ik_\alpha}+1}{e^{ik_\alpha+ik_\beta}-2e^{ik_\beta}+1}\Big].
\end{equation}
%
The eigenenergy associated to the eigenstate~\eref{ba-eig} is  
%
\begin{equation}
\label{ba-ener}
E=\sum\limits_{\alpha=1}^M(\cos(k_\alpha)-1). 
\end{equation}
%
The quasi-momenta $k_\alpha$ are obtained by solving the so-called Bethe 
equations~\cite{bethe-1931}
%
\begin{equation}
\label{ba-eq}
e^{ik_\alpha L}=\prod\limits^M_{\beta\ne\alpha}\Big[-\frac{1-2e^{
ik_\alpha}-e^{ik_\alpha+ik_\beta}}{1-2e^{ik_\beta}-e^{ik_\alpha+
ik_\beta}}\Big].
\end{equation}
%
It is useful to  introduce the rapidities $\{\lambda_\alpha\}_{\alpha=1}^M$ as 
%
\begin{equation}
\label{rap}
k_\alpha=\pi-2\arctan(\lambda_\alpha)\quad\mbox{mod}\, 2\pi.
\end{equation}
%
Taking the logarithm on both sides in~\eref{ba-eq}, and using~\eref{rap}, 
one obtains the Bethe equations in logarithmic form as 
%
\begin{equation}
\label{ba-eq-log}
\arctan(\lambda_\alpha)=\frac{\pi}{L}J_\alpha+\frac{1}{L}\sum\limits_{
\beta\ne\alpha}\arctan\Big(\frac{\lambda_\alpha-\lambda_\beta}{2}\Big),
\end{equation}
%
where $-L/2<J_\alpha\le L/2$ are the so-called Bethe quantum numbers. The 
$J_\alpha$ are half-integers and integers for $L-M$ even and odd, 
respectively.

Finally, one should remark that $M$-particle eigenstates corresponding to 
{\it finite} rapidities are eigenstates with maximum allowed magnetization 
(highest-weight eigenstates) $S_T^z=L/2-M=S_T$, with $S_T$ the total spin. 
Due to the $SU(2)$ invariance of~\eref{xxx-ham}, all the states in the same 
$S_T$ multiplet and different $-S_T\le S_T^z\le S_T$ are eigenstates of the 
$XXX$ chain with the same energy eigenvalue. These eigenstates (descendants) 
are obtained by multiple applications of the total-spin lowering operator 
$S_T^-\equiv\sum_iS_i^-$ on the highest-weight eigenstates. In the Bethe ansatz 
framework the rapidities of a generic $M$-particle descendant eigenstate with 
$S_T^T=L/2-M'$, with $M'<M$, are obtained by supplementing the $M$ rapidities 
of the highest-weight state with $M'-M$  infinite rapidities. We anticipate 
that descendant eigenstates have non-zero overlap with the zero-momentum 
N\'eel state (cf. section~\ref{sec:2}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{String hypothesis \& the Bethe-Gaudin-Takahashi (BGT) equations}
\label{sec:1.3}

In the thermodynamic limit $L\to\infty$ the solutions of the Bethe 
equations~\eref{ba-eq} form particular ``string'' patterns in the complex plane, 
(string hypothesis)~\cite{bethe-1931,taka-book}. Specifically, the rapidities 
forming a ``string'' of length $1\le n\le M$ (that we defined here as 
$n$-string) can be parametrized as 
%
\begin{equation}
\label{str-hyp}
\lambda^{j}_{n;\gamma}=\lambda_{n;\gamma}-i(n-1-2j)+i\delta_{n;\gamma}^j,\qquad 
j=0,1,\dots, n-1, 
\end{equation}
%
with $\lambda_{n;\gamma}$ being the real part of the string (string center), 
$\gamma$ labelling strings with different centers, and $j$ labelling the different 
components of the string. In Eq.~\eref{str-hyp} $\delta_{n;\gamma}^j$ are the string 
deviations, which typically, i.e., for most of the chain eigenstates, vanish 
exponentially with $L$ in the thermodynamic limit. 
Note that real rapidities correspond to strings of unit length ($1$-strings), i.e., 
$n=1$ in Eq.~\eref{str-hyp}.

The string centers $\lambda_{n;\gamma}$ are obtained by solving the so-called 
Bethe-Gaudin-Takahashi equations
%
\begin{equation}
\label{bgt-eq}
2L\theta_n(\lambda_{n;\gamma})=2\pi I_{n;\gamma}+\sum\limits_{(m,
\beta)\ne(n,\gamma)}\Theta_{m,n}(\lambda_{n;\gamma}-\lambda_{m;\beta}).  
\end{equation}
%
The generalized scattering phases $\Theta_{m,n}$ read 
%
\begin{eqnarray}
\label{Theta}
\nonumber\fl\Theta_{m,n}(x)\equiv\left\{\begin{array}{cc}
\theta_{|n-m|}(x)+\!\!\!\!\!\sum
\limits_{r=1}^{(n+m-|n-m|-1)/2}\!\!\!\!\!2\theta_{|n-m|+2r}(x)
+\theta_{n+m}(x) & \quad\mbox{if}
\quad n\ne m\\\fl\sum\limits_{r=1}^{n-1}2\theta_{2r}(x)+
\theta_{2n}(x) & \quad\mbox{if}\quad n=m
\end{array}\right.
\end{eqnarray}
%
with $\theta_\alpha(x)\equiv 2\arctan(x/\alpha)$. Here $I_{n;\gamma}$ are the 
Bethe-Takahashi quantum numbers associated with $\lambda_{n;\gamma}$. 
The solutions of~\eref{bgt-eq}, and the Bethe states thereof, are naturally 
classified according to their ``string content'' ${\mathcal S}\equiv\{s_n
\}_{n=1}^M$, with $s_n$ the number of $n$-strings. Clearly, the constraint 
$\sum_{n=1}^{M}n s_n=M$ has to be satisfied. It can be shown that 
the BGT quantum numbers $I_{n;\gamma}$ associated with the $n$-strings are 
integers and half-integers for $L-s_n$ odd and even, respectively. 
An upper bound for the BGT quantum numbers can be derived as~\cite{taka-book} 
%
\begin{equation}
|I_{n;\gamma}|\le I^{(MAX)}_{n}\equiv\frac{1}{2}(L-1-\sum
\limits_{m=1}^Mt_{m,n}s_m),
\label{bt-qn-bound}
\end{equation}
%
where $t_{m,n}\equiv 2\mbox{min}(n,m)-\delta_{m,n}$. Using the string hypothesis 
Bethe states energy eigenvalue~\eref{ba-ener} becomes
%
\begin{equation}
\label{ener-str}
E=-\sum_{n,\gamma}\frac{2n}{\lambda_{n;\gamma}^2+n^2}. 
\end{equation}
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Thermodynamic limit}
\label{sec:1.4}

In the thermodynamic limit $L\to\infty$ at fixed finite particle density $M/L$ 
the solutions of the BGT equations~\eref{bgt-eq} become dense. One then defines 
the BGT root distributions for the $n$-strings as $\pmb{\rho}\equiv\{\rho_n(
\lambda)\}_{n=1}^\infty$, with $\rho_n(\lambda)\equiv\lim_{L\to\infty}[
\lambda_{n;\gamma+1}-\lambda_{n;\gamma}]^{-1}$. Thus the  BGT 
equations~\eref{bgt-eq} become an infinite set of coupled non-linear integral 
equations for the $\rho_n(\lambda)$ as 
%
\begin{equation}
\label{bgt-th}
a_n(\lambda)=\rho_n(\lambda)+\rho^h_n(\lambda)+\sum_m(T_{n,m}*\rho_m)
(\lambda),
\end{equation}
%
where $\rho_n^{h}$ are the hole-distributions, and the functions $a_n(\lambda)$ 
are defined as 
%
\begin{equation}
a_n(x)\equiv\frac{1}{\pi}\frac{n}{x^2+n^2}. 
\end{equation}
%
In~\eref{bgt-th} $T_{n,m}*\rho_m$ denotes the convolution 
%
\begin{equation}
(T_{n,m}*\rho_m)(\lambda)\equiv\int_{-\infty}^{+\infty}T_{n,m}(\lambda-\lambda')
\rho_{m}(\lambda'),
\end{equation}
%
with the matrix $T_{n,m}(x)\equiv\Theta'(x)$ being dfined as 
%
\begin{eqnarray}
\nonumber\fl T_{m,n}(x)\equiv\left\{\begin{array}{cc}
a_{|n-m|}(x)+\!\!\!\!\!\sum
\limits_{r=1}^{(n+m-|n-m|-1)/2}\!\!\!\!\!2a_{|n-m|+2r}(x)
+a_{n+m}(x) & \quad\mbox{if}
\quad n\ne m\\\fl\sum\limits_{r=1}^{n-1}2a_{2r}(x)+
a_{2n}(x) & \quad\mbox{if}\quad n=m
\end{array}\right.
\end{eqnarray}
%
For a generic smooth enough observable ${\mathcal O}$ in the thermodynamic 
limtit its expectation value is expected to become a functional of the root 
densities $\pmb{\rho}$ as $\langle\pmb{\rho}|{\mathcal O}|\pmb{\rho}\rangle$.

Moreover, for all the local observables considered in this work the contribution of the 
different type of strings factorize and one can write 
%
\begin{equation}
\langle\pmb{\rho}|{\mathcal O}|\pmb{\rho}\rangle=\sum_{n=1}^\infty
\int_{-\infty}^{+\infty}d\lambda \rho_n(\lambda) {\mathcal O}_n(\lambda), 
\end{equation}
% 
with ${\mathcal O}_n(\lambda)$ the ... .

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The conserved charges}
\label{sec:1.5}

The $XXX$ chain exhibits an extensive number of mutually commuting local conserved 
charges $Q_n$, with $n\in\mathbb{N}$, i.e., 
%
\begin{equation}
[Q_n,{\mathcal H}]=0\quad\forall n\quad\textrm{and}\quad [Q_n,Q_m]=0\quad\forall n,m.
\end{equation}
%
The eigenvalues $q_n$ of the conserved charges over the Bethe eigenstates 
(cf.~\eref{ba-eig}) are given as 
%
\begin{equation}
\label{Q-def}
\left.q_{n+1}\equiv\frac{i}{(n-1)!}\frac{d^n}{dy^n}\log\tau
(y)\right|_{y=i},
\end{equation}
%
where $y$ is a spectral parameter and $\tau(y)$ is the Bethe state eigenvalue of the 
so-called transfer matrix in the Algebraic Bethe Ansatz framework~\cite{kor-book}. 
The analytic expression for $\tau(y)$ in terms of the solutions 
$\{\lambda_\alpha\}_{\alpha=1}^M$ of the Bethe equations~\eref{ba-eq} is given as 
%
\begin{equation}
\label{tau}
\tau(y)\equiv\Big(\frac{y+i}{2}\Big)^L\prod\limits_\alpha\frac{y-\lambda_\alpha-2i}
{y-\lambda_\alpha}+\Big(\frac{y-i}{2}\Big)^L\prod\limits_\alpha\frac{y-\lambda_\alpha
+2i}{y-\lambda_\alpha}.
\end{equation}
%
Interestingly, from~\eref{Q-def} one has that the second term in~\eref{tau} does not 
contribute to $q_n$, at least for $n\le L-2$. 
For a generic Bethe state $q_n$ is obtained by summing independently the contributions 
of the BGT roots as 
%
\begin{equation}
\label{qngnk}
q_n=\sum_{k,\gamma}g_{n,k}(\lambda_{k;\gamma}).
\end{equation}
%
Using the string hypothesis~\eref{str-hyp}, and~\eref{Q-def},~\eref{tau}, one obtains  
the the first few functions $g_{n,k}$ in terms of the solutions of the BGT equations as 
%
\begin{eqnarray}
\label{gnk}
g_{3,k}=-\frac{4k\lambda_{k;\gamma}}{(\lambda_{k;\gamma}^2+k^2)^2}, &\quad  
g_{4,k}=\frac{2k(k^2-3\lambda_{k;\gamma}^2)}{(k^2+\lambda_{k;\gamma}^2)^3}\\\nonumber 
g_{5,k}=\frac{8k\lambda_{k;\gamma}(k^2-\lambda_{k;\gamma}^2)}{(k^2+
\lambda_{k;\gamma}^2)^4}, & \quad
g_{6,k}=-\frac{2k(5\lambda_{k;\gamma}^4-10k^2\lambda_{k;\gamma}^2+k^4)}{(k^2+
\lambda_{k;\gamma}^2)^5}. 
\end{eqnarray}
%
Note that $q_2$ is the Bethe state energy eigenvalue and is given in~\eref{ener-str}. 
It is also interesting to observe that $g_{n,k}$ (cf.~\eref{gnk}) is vanishing for 
infinite BGT roots. This is expected to hold for the generic $g_{n,k}$, i.e., for any 
$n$, and it is a consequence of the $SU(2)$ invariance of the conserved charges. Finally, 
in the thermodynamic limit $L\to\infty$ from~\eref{qngnk} one has 
%
\begin{equation}
\label{q0-th}
q_n\to\sum_{k=1}^\infty\int_{-\infty}^{+\infty}d\lambda\rho_k(\lambda)g_{n,k}(\lambda), 
\end{equation}
%
where the BGT root distributions $\rho_k(\lambda)$ are solutions of~\eref{bgt-th}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overlap between the Bethe states and some initial states} 
\label{sec:2}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{N\'eel state overlap}
\label{sec:2.1}

Here we detail the Bethe ansatz results for the overlap of the Bethe states (cf.~\eref{ba-eig}) 
with the zero-momentum (one-site shift invariant) N\'eel state $|N\rangle$ and the 
Majumdar-Ghosh (MG) $|MG\rangle$ state. We stard discussing the N\'eel state. This is 
defined as 
%
\begin{equation}
\label{neel}
|N\rangle\equiv\frac{1}{\sqrt{2}}\big(|N_1\rangle
+|N_2\rangle\big), 
\end{equation}
%
with $|N_1\rangle\equiv |\uparrow\downarrow\rangle^{\otimes L/2}$, and $|N_2\rangle\equiv 
|\downarrow\uparrow\rangle^{\otimes L/2}$. Note that $|N_1\rangle=\hat T|N_2\rangle$, 
with $\hat T$ the one-site shift operator.

Due to the zero-momentum constraint, only parity-invariant Bethe states can have non-zero 
N\'eel overlap~\cite{brockmann-2014}. Parity-invariant Bethe states contain only pairs of 
solutions of the Bethe equations~\eref{ba-eq} with opposite sign.  Here we denote the generic 
parity-invariant rapidity configuration as $|\{\pm\tilde\lambda_\alpha\}_{\alpha=1}^m,n_\infty
\rangle$, i.e., considering only positive rapidities. Here $m$ is the number of rapidity pairs. 
Since the N\'eel state is not invariant under $SU(2)$ rotations, eigenstates with 
infinite rapidities can have non-zero N\'eel overlaps.
Here the number of infinite rapidities is denoted as $N_{\infty}$. Note that one has $M=L/2=
N_\infty+2m$. We denote as $n_\infty\equiv N_\infty/L$ the density of infinite rapidities. 
Finally, the overlap between the Bethe states and the Neel state $|N\rangle$ 
reads~\cite{brockmann-2014,pozsgay-2014} 
%
\begin{equation}
\label{neel-ov}
\frac{\langle N|\{\pm\tilde\lambda_j\}_{j=1}^m,n_\infty\rangle}{|||\{\tilde\lambda_j\}_{j=1}^m,
n_\infty\rangle||}=\frac{\sqrt{2}N_{\infty}!}{\sqrt{(2N_\infty)!}}\left[\prod_{j=1}^m
\frac{\sqrt{\tilde\lambda_j^2+1}}{4\tilde\lambda_j}\right]\sqrt{\frac{\textrm{det}_m(G^+)}{
\textrm{det}_m(G^-)}}.
\end{equation}
%
The matrix $G^\pm$ is  defined as  
%
\begin{equation}
\label{G-pm}
\fl\quad G^{\pm}_{jk}=\delta_{jk}\Big(LK_{1/2}(\tilde\lambda_j)-\sum\limits_{l=1}^mK_1^+(
\tilde\lambda_j,\tilde\lambda_l)\Big)+K_{1}^{\pm}(\tilde\lambda_j,\tilde\lambda_k),
\quad\,j,k=1,\dots,m, 
\end{equation}
%
where 
%
\begin{equation}
\label{K}
K_1^\pm(\lambda,\mu)=K_1(\lambda-\mu)\pm K_1(\lambda+\mu) \quad\textrm{with}\quad 
K_\alpha(\lambda)\equiv\frac{8\alpha}{\lambda^2+4\alpha^2}. 
\end{equation}
%
Note that our definitions of $K_{\alpha}(\lambda)$ differs from the one in 
Ref.~\cite{brockmann-2014}, due to a factor $2$ in the definition of the 
rapidities (see also~\eref{str-hyp}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The string hypothesis: Reduced N\'eel overlap formulas}
\label{sec:2.2}

Here we consider the overlap formula for the Neel state~\eref{neel-ov} in the 
limit $L\to\infty$, assuming that the rapidities form perfect strings, i.e., 
$\delta_{n;\gamma}^j=0$ in~\eref{str-hyp}.  
Then it is possible to rewrite~\eref{neel-ov} in terms of the string centers 
$\tilde\lambda_{n;\alpha}$ only. Here we restrict ourselves to rapidity 
configurations with no zero-momentum strings, i.e., with finite string 
centers. Our results are not valid for zero-momentum strings.

First, we restrict ourselves to parity-invariant string configurations, 
i.e., considering only strings having positve string centers. We denote the 
generic parity-invariant string configuration as $\{\tilde\lambda_{n;
\gamma}\}$, where $\gamma$ labels the different non-zero string centers, and 
$n$ is the string length.

It is convenient to first split the indices $i,j$ of $G^\pm_{ij}$ as $i=(n,\gamma,i)$ 
and $j=(m,\gamma',j)$, with $n,m$ being the length of the strings, $\gamma,\gamma'$ 
labelling the corresponding string centers, and $i,j$ the components of the 
two strings.

Using~\eref{G-pm} and~\eref{K}, one has that for two consecutive rapidities in the same 
string, i.e., for $m=n,\gamma=\gamma',|i-j|=2$, the matrices $G^{\pm}_{jk}$ become 
ill-defined in the thermodynamic limit. Precisely, since $K_{1}(\tilde\lambda_{n;\gamma}^i-
\tilde\lambda_{n;\gamma}^{i+1})\sim 1/(\delta_{n;\gamma}^i-\delta_{n;\gamma}^{i+1})$, 
$G_{ij}^\pm$ diverges in the thermodynamic limit. Importantly, as the same type of 
divergence occur in both $G^+$ and $G^-$, their ratio (cf.~\eref{neel-ov}), and 
the overlaps, are finite.

The finite part of the overlaps~\eref{neel-ov} can be calculated using the same strategy 
as in Ref.~\cite{calabrese-2007,calabrese-2007-a} (see also Ref.~\cite{brockmann-2014}). 
The resulting matrix $G^+$ depends only on the ``string center'' indices $(n,\gamma)$ 
and $(m,\gamma')$ and it is given as  
%
\begin{equation}
\label{red-G+}
\fl \frac{1}{2}G^+_{(n,\gamma)(m,\gamma')}=\left\{\begin{array}{cc}
L\theta_n'(\tilde\lambda_{n;\gamma}) -\sum\limits_{(\ell,\alpha)\ne(n,\gamma)}\Big[\Theta'_{n,\ell}
(\tilde\lambda_{n;\gamma}-\tilde\lambda_{\ell;\alpha}) & \quad\textrm{if}\,(n,\gamma)=(m,\gamma') \\
+\Theta'_{n,\ell}(\tilde\lambda_{n;\gamma}+\tilde\lambda_{\ell;\alpha})\Big] & \\ \\
\Theta'_{n,m}
(\tilde\lambda_{n;\gamma}-\tilde\lambda_{m;\gamma'})+\Theta'_{n,m}
(\tilde\lambda_{n;\gamma}+\tilde\lambda_{m;\gamma'}) & \quad\textrm{if}\,(n,\gamma)\ne(m,\gamma')
\end{array}\right.
\end{equation}
%
Here $\theta_n'(x)\equiv d\theta_n(x)/dx=2n/(n^2+x^2)$ and $\Theta'(x)\equiv d\Theta(x)/dx$, 
with $\Theta(x)$ as defined in~\eref{Theta}. 
Similarly, for $G^-$ one obtains 
%
\begin{equation}
\label{red-G-}
\fl\frac{1}{2}G^-_{(n,\gamma)(m,\gamma')}=\left\{\begin{array}{cc}
\fl(L-1)\theta'_n(\tilde\lambda_{n;\gamma})-2\sum\limits_{k=1}^{n-1}\theta'_k(
\tilde\lambda_{n;\gamma})
& \textrm{if}\,(n,\gamma)= (m,\gamma')\\
-\hspace{-.5cm}\sum\limits_{(\ell,\alpha)\ne(n,\gamma)}\Big[\Theta'_{n,\ell}
(\tilde\lambda_{n;\gamma}-\tilde\lambda_{\ell;\alpha})+\Theta'_{n,\ell}
(\tilde\lambda_{n;\gamma}+\tilde\lambda_{\ell;\alpha})\Big] \\\\
\Theta'_{n,m}
(\tilde\lambda_{n;\gamma}-\tilde\lambda_{m;\gamma'})-\Theta'_{n,m}
(\tilde\lambda_{n;\gamma'}+\tilde\lambda_{m;\gamma'})) & \textrm{if}\,(n,\gamma)\ne(m,\gamma')
\end{array}\right.
\end{equation}
%
Note that in presence of zero-momentum strings, additional divergences as $1/(\delta_{n;
\gamma}^{i}+\delta_{n;\gamma}^{i+1})$ appear due to the term $K_1(\lambda+\mu)$ 
(cf.~\eref{K}). It turns out that the treatment of these diverges is a challenging 
task because it requires the precise knowledge of the string deviations, meaning 
their dependence on $L$, for each different type of string. Some results have been 
provided for small strings in Ref.~\cite{calabrese-2014}.

Finally, using the string hypothesis and the parity-invariance condition, the multiplicative 
prefactor in~\eref{neel-ov} can be simplified. Here we focus on a generic eigenstate of 
the $XXX$ identified by $m$ pairs of finite rapidities. 
Note that due to parity invariance and the exclusion of zero-momentum strings, only 
strings of length up to $m$ are allowed. The ``reduced'' string content 
identifying the eigenstate is denoted as $\widetilde{\mathcal S}=\{\tilde s_1,\dots,\tilde s_{m}\}$, 
with $\tilde s_n$ the number of parity-invariant pairs of $n$-strings. Using the string 
hypothesis and~\eref{neel-ov}, one can write 
%
\begin{equation}
\label{neel-k}
\fl\quad\prod\limits_{j=1}^m\frac{\sqrt{\tilde\lambda_j^2+1}}{4\tilde\lambda_j}=
\frac{1}{4^m}\prod\limits_{j=1}^m\prod\limits_{\ell=1}^{\tilde s_j}\left[\frac{\sqrt{j^2+
\lambda^2_{j;\ell}}}{\lambda_{j;\ell}}
\prod\limits_{k=0}^{\lceil j/2\rceil-1}\frac{(2k)^2+\lambda^2_{j;\ell}}{(2k+1)^2+
\lambda^2_{j;\ell}}\right]^{(-1)^j}.
\end{equation}
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Overlap with the Majumdar-Ghosh state}
\label{sec:2.3}

The Majumdar-Ghosh state $|MG\rangle$ is defined as 
%
\begin{equation}
\label{mg}
|MG\rangle\equiv \Big(\frac{|\uparrow\downarrow\rangle-|\downarrow\uparrow\rangle}
{\sqrt{2}}\Big)^{\otimes L/2}. 
\end{equation}
%
The overlap between a generic eigenstate of the $XXX$ chain and the Majumdar-Ghosh 
state can be obtained from the N\'eel state overlap in Eq.~\eref{neel-ov} 
as~\cite{pozsgay-2014} 
%
\begin{equation}
\label{mg-ov}
\langle MG|\{\pm\lambda_j\}_{j=1}^m\rangle=\prod\limits_{j=1}^m\frac{1}{2}
\Big(1-\frac{\lambda_j-i}{\lambda_j+i}\Big)
\Big(1+\frac{\lambda_j+i}{\lambda_j-i}\Big)
\langle N|\{\pm\lambda_j\}_{j=1}^m\rangle
\end{equation}
%
The mutliplicative factor in~\eref{mg-ov}, using the string hypothesis is 
rewritten as 
%
\begin{eqnarray}
\label{mg-k}
\fl\prod\limits_{j=1}^m\frac{1}{2}
\Big(1-\frac{\lambda_j-i}{\lambda_j+i}\Big)
\Big(1+\frac{\lambda_j+i}{\lambda_j-i}\Big)=\\\nonumber\qquad
2^m\prod\limits_{j=1}^m\prod\limits_{\ell=1}^{\tilde s_j}
\lambda_{j;\ell}^{1+(-1)^j}(\lambda^2_{j;\ell}+j^2)\prod\limits_{k=0}^{\lfloor 
j/2\rfloor}\Big[\lambda_{j;\ell}^2+\Big(2k+\frac{1-(-1)^j}{2}\Big)^2\Big]^{-2}
\end{eqnarray}
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The N\'eel overlap in the thermodynamic limit}
\label{sec:2.4}

In the thermodynamic limit $L\to\infty$ the expression for the extensive part of the 
overlap~\eref{neel-ov} can be written as~\cite{brockmann-2014} 
%
\begin{equation}
\label{neel-ov-th}
\fl-\lim_{L\to\infty}\log\left[\frac{\langle N|\{\pm\lambda_j\}_{j=1}^m,n_\infty
\rangle}{|||\{\lambda_j\}_{j=1}^m,n_\infty\rangle||}\right]=\frac{L}{2}\Big(
n_\infty\log2+\sum_{n=1}^\infty\int_0^\infty d\lambda\rho_n(\lambda)[g_n(\lambda)
+2n\log(4)]\Big), 
\end{equation}
%
where  
%
\begin{equation}
\fl\quad g_n(\lambda)=\sum_{l=1}^{n-1}\Big[f_{n-1-2l}(\lambda)-f_{n-2l}(\lambda)
\Big],\quad\textrm{with}\quad f_n(\lambda)=\log\Big(\lambda^2+\frac{n^2}{4}\Big),
\end{equation}
and
\begin{equation}
n_\infty=1-2\sum_{m=1}^\infty m\int_{-\infty}^\infty d\lambda\rho_m(\lambda). 
\end{equation}
%
Note that~\eref{neel-ov-th} is extensive. Specifically, in Eq.~\eref{neel-ov-th} 
subleading (i.e., subextensive) contributions originating from the determinant
ratio $\det_m(G^+)/\det_m(G^-)$ are neglected. Moreover, the term in~\eref{neel-ov-th} 
acts as a driving term in the Quech Action formalism (cf. section~\ref{sec:4}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Quench Action treatment of the steady state}
\label{sec:4}

In the thermodynamic limit the sum over the model eigenstates in~\eref{d-ensemble} 
can be recast into a functional integral over the BGT root distributions 
$\pmb{\rho}\equiv\{\rho_n(\lambda)\}_{n=1}^\infty$ (cf. section~\ref{sec:1.4}) 
as
%
\begin{equation}
\label{eig-sum}
\sum\limits_{\alpha}\rightarrow\int{\mathcal D}\pmb{\rho} e^{S_{YY}(\pmb{\rho})}. 
\end{equation}
%
Here ${\mathcal D}\pmb{\rho}\equiv\prod_{n=1}^\infty{\mathcal D}\rho_n(\lambda)$, 
$\rho_n(\lambda)$, and $S_{YY}(\pmb{\rho})$ is the Yang-Yang entropy, which counts 
the number of Bethe states leading to the same $\pmb{\rho}$ in the thermodynamic 
limit. Using~\eref{eig-sum}, for a generic observable ${\mathcal O}$, its diagonal 
ensemble expectation value~\eref{d-ensemble} becomes 
%
\begin{equation}
\label{qa-d-ensemble}
\quad\langle{\mathcal O}\rangle=\int{\mathcal D}\pmb{\rho}\exp\Big[2\Re\log\langle
\Psi_0|\pmb{\rho}\rangle +S_{YY}(\pmb{\rho})\Big]\langle\pmb{\rho}|{\mathcal O}|
\pmb{\rho}\rangle
\end{equation}
%
Here it is assumed that in the thermodynamic limit the eigenstate expectation 
values $\langle\alpha|{\mathcal O}|\alpha\rangle$ (cf.~\eref{d-ensemble}) 
and the overlaps $\langle\Psi_0|\pmb{\rho}\rangle$ become smooth functionals 
of the root distributions $\pmb{\rho}$.

The functional integral in~\eref{qa-d-ensemble} can be evaluated in the limit 
$L\to\infty$ using the saddle point approximation. One has to minimize the 
functional ${\mathcal F}(\pmb{\rho})$ defined as 
%
\begin{equation}
L{\mathcal F}(\pmb{\rho})\equiv 2\Re\log\langle\pmb{\rho}|\Psi_0
\rangle+S_{YY}(\pmb{\rho}(\lambda))  
\end{equation}
% 
with respect to $\pmb{\rho}$, i.e., solving $\delta{\mathcal F}(\pmb{\rho})/\delta
\pmb{\rho}|_{\pmb{\rho}=\pmb{\rho^*}}=0$, under the constraint that the thermodynamic BGT 
equations~\eref{bgt-th} hold. Finally, one obtains from~\eref{qa-d-ensemble} that in the 
thermodynamic limit 
%
\begin{equation}
\label{obs-th}
\langle{\mathcal O}\rangle=\langle\pmb{\rho^*}|{\mathcal O}|\pmb{\rho^*}\rangle. 
\end{equation}
%
Remarkably, for the quench with initial state the N\'eel state $|\Psi_0\rangle=|N\rangle$ the 
saddle point root distributions ${\rho_n^*(\lambda)}_{n=1}^\infty$ can be obtained 
analytically~\cite{brockmann-2014}. The first few are given as 
%
\begin{eqnarray}
\label{rho1-sp}
\fl\quad\rho^*_1(\lambda)=\frac{8(4+\lambda^2)}{\pi(19+3\lambda^2)(1+6\lambda^2+
\lambda^4)}\\\label{rho2-sp}
\fl\quad\rho_2^*(\lambda)=\frac{8\lambda^2(9+\lambda^2)(4+3\lambda^2)}{\pi(2+\lambda^2)
(16+14\lambda^2+\lambda^4)(256+132\lambda^2+9\lambda^4)}\\\label{rho3-sp}
\fl\quad\rho_3^*(\lambda)=\frac{8(1+\lambda^2)^2(5+\lambda^2)(16+\lambda^2)(21+\lambda^2)}
{\pi(19+3\lambda^2)(9+624\lambda^2+262\lambda^4+32\lambda^6+\lambda^8)
(509+5\lambda^2(26+\lambda^2))}.
\end{eqnarray}
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{N\'eel overlaps: The role of the zero-momentum strings Bethe states}
\label{sec:5}

In this section we discuss generic features of the overlaps between the 
eigenstates (Bethe states) of the Heisenberg spin chain and the N\'eel state 
(cf.~\eref{neel}). We exploit the Bethe ansatz solution of the chain (see 
section~\ref{sec:1}) as well as exact results for the N\'eel overlaps (see 
section~\ref{sec:2}). We focus on finite chains with $L\lesssim 40$ sites, 
considering, for any $L$, the full Hilbert space of the chain. 
The Bethe states having, in principle, non zero N\'eel overlap are the 
so-called parity-invariant Bethe states (see section~\ref{sec:2}). We 
denote their total number as $Z_{Neel}$. Crucially, here we restrict 
ourselves to the subset of parity-invariant Bethe states corresponding 
to solutions of the Bethe-Gaudin-Takahashi (BGT) equations~\eref{bgt-eq} 
that do not contain zero-momentum strings. We denote the total number of 
these eigenstates as $\widetilde Z_{Neel}$. Both $Z_{Neel}$ and $\widetilde 
Z_{Neel}$ are given in terms of the chain length $L$ by simple combinatorial 
formulas that we provide.

%##################################################################
\begin{figure}[t]
\begin{center}
\includegraphics[width=.9\textwidth]{./draft_figs/Neel_overlaps}
\end{center}
\caption{ N\'eel overlaps with the eigenstates of the Heisenberg spin 
 chain: Numerical results obtained from the full scanning of the chain 
 Hilbert space. Eigenstates corresponding to zero-momentum strings are 
 escluded. (a) Squared overlaps $|\langle\lambda|N\rangle|^2$ plotted as 
 function of the eigenstates energy density $E/L$. Here $|\lambda\rangle$ 
 denotes the generic eigenstate. The data are for chains with length $26
 \le L\le 38$. The inset is to highlight the exponential decay as a 
 function of $E/L$. Note the logarithmic scale on the $y$ axis. (b) 
 Overlap distribution function: Histograms of $-2\log\langle\lambda|N\rangle/L$. 
 The $y$-axis is rescaled by a factor $10^5$ for convenience. The dash-dotted 
 and dashed vertical lines are the N\'eel overlaps with the $XXX$ chain ground 
 state and the ferromagnetic state, respectively. The dotted line (see 
 the arrow) is the result obtained using the quench action approach. 
}
\label{fig0:neel-ov}
\end{figure}
%##################################################################

Interestingly, the fraction of eigenstates with no zero-momentum strings, i.e.,  
$\widetilde Z_{Neel}/Z_{Neel}$, is vanishing as $L^{-1/2}$ in the thermodynamic 
limit, meaning that zero-momentum strings eigenstates are dominant, at least 
in number, for large chains. This has dramatic consequences, for instance, at 
the level of the overlap sum rules for the local conservation laws of the model. 
Precisely, although the sum rules are fixed by the initial (N\'eel) state 
expectation value of the local charges, we observe striking violations for 
any finite chain. Moreover, all the sum rules exhibit vanishing behavior as 
$L^{-1/2}$ upon increasing the chain size, reflecting the same vanishing behavior 
as $\widetilde Z_{Neel}/Z_{Neel}$. A similar scenario holds for the overlaps with 
the Majumdar-Ghosh state, where, however, excluding the zero-momentum strings leads 
to a $1/L$ behavior. This demonstrates that for finite chains the physical effects 
of zero-momentum strings eigenstates are not negligible.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{N\'eel overlap distribution function: Overview}
\label{sec:5.1}

Here we overview the Bethe ansatz results for the N\'eel overlaps with the 
eigenstates of the $XXX$ chain. The total number of partiy-invariant 
eigenstates $Z_{Neel}$ having, in principle, non-zero N\'eel overlap is given 
as 
%
\begin{equation}
\label{zneel1}
Z_{Neel}=2^{\frac{L}{2}-1}+\frac{1}{2}B\Big(\frac{L}{2},\frac{L}{4}\Big)+1, 
\end{equation}
%
with $B(n,m)\equiv n!/(m!(n-m)!)$ the Newton binomial. The proof of~\eref{zneel1} 
is obtained by counting all the parity-invariant BGT quantum number configurations, 
and it is reported in~\ref{app-1}. Note that $Z_{Neel}$ provides only an upper 
bound for the number of eigenstates with non-zero N\'eel overlap, as it is clear 
from the exact diagonalization results shown in Table~\ref{table:neel}. This is 
because parity-invariant eigenstates with a single zero-momentum even-length 
string, which are included in~\eref{zneel1}, have identically zero N\'eel 
overlap~\cite{brockmann-2014}. This is not related to the symmetries of the N\'eel 
state, but to the ``accidental'' vanishing of the prefactor in the overlap 
formula~\eref{neel-ov}. Finally, after excluding the zero-momentum strings 
eigenstates, the total number of remaining eigenstates $\widetilde Z_{Neel}$, 
which are the ones considered here, is given as (see~\ref{app-2} for the proof) 
%
\begin{equation}
\label{ztilde}
\widetilde Z_{Neel}=B\Big(\frac{L}{2},\frac{L}{4}\Big).
\end{equation}
%
An overview of generic features of the overlaps is given in Figure~\ref{fig0:neel-ov} 
(a) plotting the squared N\'eel overlaps $|\langle\lambda|N\rangle|^2$ with the $XXX$ 
chain eigenstates $|\lambda\rangle$ versus the energy density $E/L$. The figure shows 
results for chains with $26\le L\le 38$ sites. The data are obtained by generating all 
the relevant parity-invariant BGT quantum numbers, and solving the associated BGT 
equations~\eref{bgt-eq}, to obtain the rapidities of $XXX$ chain eigenstates. The 
N\'eel overlaps are calculated numerically using~\eref{neel-ov}. Note that for $L=38$ 
from~\eref{ztilde} the total number of overlap shown in the Figure is $\widetilde Z_{Neel}
\sim 10^5$.

Clearly, the overlaps decay exponentially as a function of $L$, as expected. Moreover, 
at each finite $L$ a rapid decay as a function of $E/L$ is observed. The inset of 
Figure~\ref{fig0:neel-ov} (a) suggests that the decay is exponential. Complementary 
information about the overlaps is shown in Figure~\ref{fig0:neel-ov} (b) plotting the 
histograms of $\kappa\equiv-2\log|\langle\lambda|N\rangle|/L$ (overlap distribution 
function). Larger values of $\kappa$, correspond to a faster decay with $L$ of the 
overlaps. The factor $1/L$ in the definition takes into account that the N\'eel 
overlaps vanish exponentially as $|\langle\lambda|N\rangle|^2\propto e^{-\kappa L}$ in 
the thermodynamic limit. Note that $\kappa$ is the driving term in the quench action 
approach (cf~\eref{neel-ov-th}). As expected, from Figure~\ref{fig0:neel-ov} (b) one 
has that the majority of the $XXX$ chain eigenstates give small N\'eel overlap (note 
the maximum at $\kappa\sim 0.5$). Interestingly, the data suggest that with 
$0.18\lesssim\kappa\lesssim0.7$. The vertical dash-dotted line in the figure is 
the $\kappa$ obtained from the N\'eel overlap of the ground state of the $XXX$ 
chain in the thermodynamic limit. This is derived using the ground state root 
distribution $\rho_1(\lambda)\propto1/\cosh(\pi\lambda)$ and~\eref{neel-ov-th}. 
On the other hand, the vertical dashed line denotes the N\'eel overlap $\sim 2/B(L,L/2)$ 
of the $S_z=0$ component of the ferromagnetic multiplet, which is at the top of the
chain energy spectrum. Finally, the vertical dotted line in Figure~\ref{fig0:neel-ov} 
(b) shows the quench action result for $\kappa$ in the thermodynamic limit. This is 
obtained by using~\eref{neel-ov-th} and the saddle point root distributions $\rho_n^*$ 
(cf.~\eref{rho1-sp}-\eref{rho3-sp} for the results up to $n=3$). Note that $\kappa$ 
does not coincide with the peak of the overlap distribution function, as expected, 
due to the effect of the driving term~\eref{neel-ov-th} in the quench action approach.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{N\'eel overlap sum rules}
\label{sec:5.2}

%##################################################################
\begin{figure}[t]
\begin{center}
\includegraphics[width=.9\textwidth]{./draft_figs/Neel}
\end{center}
\caption{Overlap sum rules for the N\'eel state $|N\rangle$: The role of 
 the zero-momentum strings. (a) The overlap sum rule $\sum_{\lambda}|
 \langle\lambda|N\rangle|^2=1$. Here $|\lambda\rangle$ are the eigenstates  
 of the $XXX$ chain. The $x$-axis shows the inverse chain length $1/L$. 
 The circles are Bethe ansatz results for chains up to $L=38$. The data  
 are obtained by  a full scanning of the chain Hilbert space. Eigenstates  
 corresponding to zero-momentum strings are excluded. The dotted line is 
 the expected result at any $L$. The data are compatible with a vanishing 
 behavior in the thermodynamic limit. The dash-dotted line is a fit to 
 $A/L^{1/2}+B/L$, with $A,B$ fitting parameters. Inset: The same data as 
 in the main Figure now plotted versus $1/L^{1/2}$. (b) 
 The same as in (a) for the energy sum rule $\sum_{\lambda}|\langle
 \lambda|N\rangle|^2Q_2(\lambda)=Q_2^{(0)}$, with $Q_2(\lambda)$ the 
 energy of the eigenstate $|\lambda\rangle$ and $Q_2^{(0)}/L=-1/2$ the 
 N\'eel state energy density (dotted line in the Figure). 
}
\label{fig1:neel-sr}
\end{figure}
%##################################################################

Here we illustrate the effect of the zero-momentum strings eigenstates on the 
N\'eel overlap sum rules. We focus on the ``trivial'' sum rule, i.e., the 
normalization of the N\'eel state  
%
\begin{equation}
\label{sr-trivial}
\langle N|N\rangle=\sum\limits_{\lambda}|\langle\lambda|N\rangle|^2=1. 
\end{equation}
%
We also consider the N\'eel expectation value of the local conserved charge 
$Q_n$ of the $XXX$ chain (see subsection~\ref{sec:1.5}), which provide 
the additional sum rules
%
\begin{equation}
\label{sr-charge}
Q_n^{(0)}=\langle N|Q_n|N\rangle=\sum\limits_{\lambda}|\langle\lambda|N\rangle|^2
Q_{n}(\lambda)\quad\textrm{with}\quad n\in\mathbb{N}, 
\end{equation}
%
where $Q_n(\lambda)$ are the charges eigenvalues over the generic Bethe state 
$|\lambda\rangle$ (cf.~\eref{qngnk} and~\eref{gnk}). 
In both~\eref{sr-trivial} and~\eref{sr-charge} the sum are restricted to the 
eigenstates with containing no zero-momentum strings. 
In~\eref{sr-charge} $Q_n^{(0)}$ is the charge expectation value over the initial 
state, i.e., the N\'eel state, which has been calculated in Ref.~\cite{fagotti-2013}. 
Due to the locality of $Q_n$, the translational invariance of the initial 
state, and the periodic boundary conditions, the density $Q_n^{(0)}/L$ does 
not depend on the chain size. For the N\'eel state, $Q_n^{(0)}$ can be calculated 
directly in the thermodynamic limit using~\eref{q0-th} and the root distributions 
$\pmb{\rho}^*$ (cf.~\eref{rho1-sp}-\eref{rho3-sp}).

The sum rules~\eref{sr-trivial} and~\eref{sr-charge} (for $n=2$, i.e., the energy 
sum rule), are shown in Figure~\ref{fig1:neel-sr} (a) and (b), respectively. 
Note that $Q^{(0)}_2/L=-1/2$ in~\eref{sr-charge} (horizonthal dotted line in 
Figure~\ref{fig1:neel-sr} (b)). 
The circles in Figure~\ref{fig1:neel-sr} (a) are the Bethe ansatz results excluding 
the zero momentum strings. The data are the same as in Figure~\ref{fig0:neel-ov}. 
The sum rules are plotted against the inverse chain length $1/L$, for $L\le 38$.

Clearly, both the sum rules are violated, due to the exclusion of the to zero-momentum 
strings. Moreover, in both Figure~\ref{fig1:neel-sr} (a) and (b) the data suggest a 
vanishing behavior upon increasing $L$. The dash-dotted lines in the Figures are fits to  
$A/L^{1/2}+B/L$, with $A,B$ fitting parameters. Interestingly, the behavior as $\propto 
L^{-1/2}$ of the sum rules reflects that of the fraction of non-zero momentum string 
eigenstates in the thermodynamic limit $\widetilde Z_{Neel}/Z_{Neel}$. Specifically, 
from~\eref{zneel1} and~\eref{ztilde} it is straightforward to derive that in the 
thermodynamic limit 
%
\begin{equation}
\label{beh}
\frac{\widetilde Z_{Neel}}{Z_{Neel}}\propto\frac{4}{\sqrt{\pi L}}. 
\end{equation}
%

%##################################################################
\begin{figure}[t]
\begin{center}
\includegraphics[width=.9\textwidth]{./draft_figs/Dimer}
\end{center}
\caption{Overlap sum rules for the Majumdar-Ghosh state $|MG\rangle$: 
 The role of the zero-momentum strings. (a) The sum rule $\sum_{\lambda}|
 \langle\lambda|MG\rangle|^2=1$, with $|\lambda\rangle$ the eigenstates   
 of the $XXX$ chain. The $x$-axis plots the inverse chain length $1/L$. 
 The circles are Bethe ansatz results for chains up to $L=32$. The results 
 are obtained by a full scanning of the chain Hilbert space. Eigenstates 
 corresponding to zero-momentum strings are excluded. The dash-dotted line 
 is a fit to $A/L+B/L^2$, with $A,B$ fitting parameters. (b) The same as 
 in (a) for the energy sum rule $\sum_{\lambda}|\langle\lambda|MG\rangle
 |^2Q_2(\lambda)=Q_2^{(0)}$, with $Q_2(\lambda)$  the energy of $|\lambda
 \rangle$ and $Q_2^{(0)}/L=-2/3$ the Majumdar-Ghosh energy density (dotted 
 line in the Figure). 
}
\label{fig2:dimer-sr}
\end{figure}
%##################################################################

It is interesting to observe that the large $L$ behavior as $L^{-1/2}$  is not 
generic, meaning that it depends on the initial state $|\Psi_0\rangle$. This is 
illustrated in Figure~\ref{fig2:dimer-sr}, focusing on the Majumdar-Ghosh (MG) 
state. As for the N\'eel state, only parity-invariant eigenstates can have 
non-zero Majumdar-Ghosh overlap. Their total number $Z_{MG}$ (cf.~\eref{p-inv-mg}) 
is given as 
%
\begin{equation}
\label{mg1}
Z_{MG}=B\Big(\frac{L}{2}-1,\frac{L}{4}-1\Big)+B\Big(\frac{L}{2}-1,
\frac{L}{4}-1\Big). 
\end{equation}
%
Note that at any size $L$ one has $Z_{MG}<Z_{Neel}$. This is due to the fact that 
the Majumdar-Ghosh state is invariant under $SU(2)$ rotations, since it contains 
only spin singlets, and it has non-zero overlap only with the $S^z_T=0$ sector 
of the $XXX$ chain spectrum. The total number of parity-invariant eigenstates 
that contain no zero-momentum strings $\widetilde Z_{MG}$ is instead (
cf.~\eref{mg-fi})
%
\begin{equation}
\label{mg2}
\widetilde Z_{MG}=B\Big(\frac{L}{2},\frac{L}{4}\Big)-B\Big(\frac{L}{2},
\frac{L}{4}-1\Big). 
\end{equation}
%
Panels (a) and (b) in Figure~\ref{fig2:dimer-sr} plot the sum rules~\eref{sr-trivial} 
and~\eref{sr-charge} for the Majumdar-Ghosh state. The data are obtained using the 
analytic results for the overlaps in subsection~\ref{sec:2.3}. The expected value for 
the energy density sum rule is $Q_2^{(0)}=-2/3$ (horizonthal dotted line in 
Figure~\ref{fig2:dimer-sr} (b)). Similar to Figure~\ref{fig1:neel-sr}, due to the exclusion 
of the zero-momentum strings, the sum rules are violated, exhibiting vanishing behavior in 
the thermodynamic limit. However, in contrast with the N\'eel case, one has the behavior 
as $1/L$, as confirmed by the fits (dash-dotted lines in Figure~\ref{fig2:dimer-sr}). 
Similar to the N\'eel case, the vanishing of the sum rules in the thermodynamic limit 
reflects the behavior of $\widetilde Z_{MG}/Z_{MG}$ as (see~\eref{mg1} and~\eref{mg2})
%
\begin{equation}
\frac{\widetilde Z_{MG}}{Z_{MG}}=\frac{4}{4+L}. 
\end{equation}
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Monte Carlo implementation of the quench action approach}
\label{sec6:mcqa}

Here, by generalizing the results presented in Ref.~\cite{alba-2015}, we present 
a Monte Carlo implementation of the quench action approach for N\'eel quench in 
the $XXX$ chain. The key idea is to sample the eigenstates of the finite-size $XXX$ 
chain with the diagonal ensemble, equivalently the quench action, probability 
distribution, given in~\eref{qa-d-ensemble}. Importantly, we consider a truncated  
Hilbert space, restricting ourselves to the eigenstates corresponding to 
solutions of the BGT equations with no zero-momentum strings. Our main physical 
result is that, despite this restriction, the remaining eigenstates contain enough 
information to correctly reproduce the post-quench thermodynamic behavior of 
the $XXX$ chain.

In subsection~\ref{sec:6.1} we detail the Monte Carlo algorithm. In 
subsection~\ref{sec:6.2} we numerically demonstrate that after the Monte Carlo 
``resampling'' the N\'eel sum rules~\eref{sr-charge} are restored, in the 
thermodynamic limit. The Hilbert space truncation is reflected only in 
$\propto 1/L$ finite-size corrections to the sum rules. In the Bethe ansatz language 
our results mean that the eigenstates sampled by the Monte Carlo approach the 
quench action representative state in the thermodynamic limit. This is explicitly 
demonstrated here by numerically extracting the quench action root distributions 
$\pmb{\rho}^*$ (cf.~\eref{rho1-sp}-\eref{rho3-sp}). The numerical results are found 
in remarkable agreement with the quench action.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The quench action Monte Carlo algorithm}
\label{sec:6.1}

The Monte Carlo procedure starts with a randomly selected parity-invariant eigenstate 
(Bethe state) of the $XXX$ chain, in the sector with zero magnetization, i.e., $M=L/2$ 
particles. Due to the fact that the N\'eel state is not invariant under $SU(2)$ rotations, 
in order to characterize the Bethe states one has to specify the number $N_{\infty}$ of 
infinite rapidities. The number of remaining particles corresponding to finite roots 
$M'$ is $M'=L/2-N_\infty$. The state is identified by a parity-invariant BGT quantum number 
configuration that we denote as ${\mathcal C}$. Due to the parity-invariance and the 
zero-momentum strings being excluded, ${\mathcal C}$ is identified by the number $m'$ 
of parity-invariant rapidities $\{\pm\lambda_j\}_{j=1}^{m'}$. The string content 
associated with the state is denoted as $\widetilde{\mathcal S}=\{\tilde s_1,\dots,
\tilde s_{m'}\}$, where $\tilde s_n$ is the number of pairs of $n$-strings. The Monte 
Carlo procedure generates a new parity-invariant eigenstate of the $XXX$ chain, and 
it consists of four steps: 
%
\begin{enumerate}
\item[\circled{1}] Choose a new number of finite-momentum particles $M''$ and of 
parity-invariant rapidity pairs $m''\equiv M''/2$ with 
probability ${\mathcal P}(M'')$ as 
%
\begin{equation}
\label{PM}
{\mathcal P}(M'')=\frac{\widetilde Z'_{Neel}(L,M'')}
{\widetilde{Z}_{Neel}(L)}, 
\end{equation}
%
where $\widetilde Z_{Neel}(L)$ is defined in~\eref{ztilde}, and $\widetilde 
Z'_{Neel}$ is the number of parity-invariant eigenstates with no zero-momentum 
strings in the sector with fixed particle number $M''$  (cf.~\eref{neel-fi} for 
the precise expression). 
\item[\circled{2}] Choose a new string content $\widetilde{\mathcal S}'\equiv
\{\tilde s_1',\dots,\tilde s'_{m''}\}$ with probability ${\mathcal P}'(M'',
\widetilde{\mathcal S}')$
%
\begin{equation}
\label{PS}
{\mathcal P}'(M'',\widetilde{\mathcal S}')=\frac{1}{\widetilde Z'_{Neel}
(L,M'')}\prod_{n=1}^{m''}B\Big(\frac{L}{2}-\sum\limits_{l=1}^{
m''}t_{nl}\tilde s'_l,\tilde s'_n\Big), 
\end{equation}
%
where the matrix $t_{nl}$ is defined in~\eref{bt-qn-bound}.
\item[\circled{3}] Generate a new parity-invariant quantum number configuration 
${\mathcal C}'$ compatible with the $\widetilde {\mathcal  S}'$ obtained in step 
$\circled{2}$. Solve the corresponding BGT equations~\eref{bgt-eq}, finding the 
rapidities $\{\pm\lambda'_j\}_{j=1}^{m''}$ of the new parity-invariant eigenstate. 
\item[\circled{4}] Calculate the N\'eel overlap $\langle\{\pm\lambda'_j\}_{j=1}^{m''}
|N\rangle$ for the new eigenstate,  using~\eref{neel-ov}~\eref{red-G+}~\eref{red-G-} 
and~\eref{neel-k}. Accept the new eigenstate with the quench action Metropolis 
probability 
%
\begin{equation}
\label{metropolis}
{\mathcal P}''_{\lambda\to\lambda'}=\textrm{Min}\Big\{1,\exp\Big(-
2\Re({\mathcal E}'-{\mathcal E})\Big)\Big\}, 
\end{equation}
%
where ${\mathcal E}'\equiv-\log\langle\{\pm\lambda'_j\}_{j=1}^{m''}
|N\rangle$, ${\mathcal E}\equiv-\log\langle\{\pm\lambda_j\}_{j=1}^{m''}
|N\rangle$, and $\Re$ denoting the real part. 
\end{enumerate}
%
Note that while the steps $1$-$3$ account for the string content and particle 
number probabilities of the parity-invariant states, step $4$ assigns to the 
different eigenstates the right quench action, equivalently diagonal ensemble, 
probability.

For a generic observable ${\mathcal O}$, its quench action, or diagonal ensemble 
average, $\langle{\mathcal O}\rangle$ is obtained from the Monte Carlo simulation 
as the arithmetic average of the eigenstates expectation values $\langle\lambda
|{\mathcal O}|\lambda\rangle$, with $|\lambda\rangle$ the eigenstates sampled by 
the Monte Carlo, as 
%
\begin{equation}
\label{qamc-obs}
\langle{\mathcal O}\rangle=\frac{1}{N_{mcs}}\sum\limits_{\lambda}\langle\lambda|
{\mathcal O}|\lambda\rangle. 
\end{equation}
%
Here $N_{mcs}$ being the total number of Monte Carlo steps. Note that, as usual in 
Monte Carlo, some initial steps have to be neglected to ensure thermalization. 
Note that~\eref{qamc-obs} can be used for any observable ${\mathcal O}$ for which 
the the Bethe state expectation value $\langle\lambda|{\mathcal O}|\lambda\rangle$ 
(form factor) is known.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The N\'eel overlap sum rules: Monte Carlo results}
\label{sec:6.2}

%##################################################################
\begin{figure}[t]
\begin{center}
\includegraphics[width=.9\textwidth]{./draft_figs/QAMC_Obs_Neel}
\end{center}
\caption{The overlap sum rules for the N\'eel state $|N\rangle$ in the 
 Heisenberg spin chain: Numerical results obtained by the Monte Carlo 
 sampling of the chain Hilbert space. In all panels the $x$ axis shows 
 the inverse chain length $1/L$ (a) The energy sum rule $\sum_\lambda|
 \langle N|\lambda\rangle|^2Q_2(\lambda)/L=Q^{(0)}_2$, with $|\lambda
 \rangle$ the generic eigenstate of the $XXX$ chain, $Q_2(\lambda)/L$ 
 the associated energy density, and $Q_2^{(0)}=-1/2$ the N\'eel energy 
 density. The symbols are Monte Carlo data. The dash-dotted line is the 
 expected result $Q_2^{(0)}$. The dashed line is a fit to the behavior 
 $-1/2+A/L+B/L^2$, with $A,B$ fitting parameters. (b) The N\'eel sum rule 
 for the energy fluctuations $\sigma^2(Q_2)$ (cf.~\eref{sigma-sr} for 
 the definition). The horizontal line is the expected result $1/4$. 
 (c)(d) Same as in (a)(b) for the charge $Q_4$ and its fluctuations. 
}
\label{fig3:neel-qamc-sr}
\end{figure}
%##################################################################

The validity of the Monte Carlo approach for simulating the quench action 
is demonstrated in Figure~\ref{fig3:neel-qamc-sr} (a)-(d). The Figure focuses on 
the N\'eel overlap sum rules for the conserved charges densities $Q_2/L$ and $Q_4/L$ 
(cf. subsection~\ref{sec:1.5} for the definition of the charges, and~\eref{sr-charge} 
for the associated sum rules). We also consider the corresponding fluctuations 
$\sigma^2(Q_n)$, which are defined as 
%
\begin{equation}
\label{sigma-sr}
\fl\quad\sigma^2(Q_n)\equiv\langle N|Q^2_n|N\rangle-\langle N|Q_n|N\rangle^2=
\sum\limits_{\lambda}|\langle N|\lambda\rangle|^2Q^2_n(\lambda)-\Big(\sum_\lambda|
\langle N|\lambda\rangle|^2Q_n(\lambda)\Big)^2.
\end{equation}
%
Note that in both~\eref{sr-charge} and~\eref{sigma-sr} the sum is now over the 
eignestates $|\lambda\rangle$ sampled by the Monte Carlo. 
Panel (a) and (b) in Figure~\ref{fig3:neel-qamc-sr} plot the sum rules for 
the energy density $Q_2/L$, and its density of fluctutations $\sigma^2(Q_2)/L$. 
It is straightforward to derive the N\'eel expectation value of the energy 
density fluctuations as $\sigma(Q_2)/L=1/4$. This is shown as dash-dotted line 
in Figure~\ref{fig3:neel-qamc-sr} (b).

The circles in the Figure are Monte Carlo data obtained using the procedure outlined 
in~\ref{sec:6.1} for the Heisenberg chain with $L\le 56$ sites. The data correspond 
to Monte Carlo simulations with $N_{mcs}\sim 10^7$ Monte Carlo steps (mcs). For the 
largest chain size $L=56$ we used $N_{mcs}\sim ...$. In all panels the $x$-axis shows 
the inverse chain length $1/L$.

Clearly, the Monte Carlo data suggest that in the thermodynamic limit the N\'eel 
overlap sum rules~\eref{sr-charge} are restored, while violations are present for 
finite chains. This numerically confirms that the truncation of the Hilbert space, 
i.e., removing the zero-momentum strings, gives rise only to scaling corrections, 
while the thermodynamic behavior after the quench is correctly reproduced. 
Note that the data in panel (a) suggest the behavior $\propto 1/L$ for the scaling 
corrections, as confirmed by the fit to $-1/2+A/L+b/L^2$ (dashed line in the Figure), 
with $A,B$ fitting parameters. The same behavior should be expected for the energy 
fluctuations $\sigma^2(Q_2)$ (panel (b) in the Figure), although the data suggest 
that this asymptotic behvior sets in for $L\gg 56$.

Similarly, panels (c) and (d) in Figure~\ref{fig3:neel-qamc-sr} plot the charge 
density $Q_4/L$ and its fluctuations $\sigma^2(Q_4)/L$. While for $Q_4/L$ the Monte 
Carlo data for $L=48$ are already compatible with the expected result $Q_4/L=1/4$ 
in the thermodynamic limit, $\sigma^2(Q_4)$ exhibits large scaling corrections. 
This could be attributed to fact that the support of $Q_n$, i.e., the number of 
sites where the operator acts non trivially, increases linearly with $n$ 
(see~\cite{grabowski-1995} for the precise expression).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Extracting the quench action root distributions}
\label{sec:6.3}

%##################################################################
\begin{figure}[t]
\begin{center}
\includegraphics[width=.95\textwidth]{./draft_figs/Neel_rho}
\end{center}
\caption{ The quench action root distributions $\rho^*_1(\lambda)$ and 
 $\rho^*_2(\lambda)$ for the $1$-strings and $2$-strings, respectively: Monte 
 Carlos results. (a) The histograms of the $1$-string Bethe-Gaudin-Takahashi 
 (BGT) roots $\lambda$ sampled in the Monte Carlo. The data are for a chain 
 with $L=56$ sites and a Monte Carlo history with $N_{mcs}\sim 10^7$ Monte 
 Carlo steps. The $y$-axis is divided by a factor $10^6$ for convenience. 
 The width of the histogram bin is $\Delta\lambda\sim 0.07$. (c) The same as 
 in (a) for the $2$-string roots. (b) The $1$-string root distribution 
 $\rho^*_1(\lambda)$ plotted versus $\lambda$ for two chains with $L=48$ 
 and $L=56$ (diamond and circles, respectively). The full line 
 is the quench action analytic result in the thermodynamic limit. (d) 
 The same as in (b) for the $2$-string root distribution $\rho^*_2(\lambda)$. 
 In both (b) and (d) the oscillations are finite-size effects, whereas 
 the error bars are the statistical Monte Carlo errors. 
}
\label{fig4:neel-rho}
\end{figure}
%##################################################################

The BGT root distributions corresponding to the quench action steady state 
(cf.~\eref{rho1-sp}-\eref{rho3-sp}) $\pmb{\rho^*}=\{\rho^*_n(\lambda)\}_{n=
1}^{\infty}$ can be extracted from the Monte Carlo simulation,  as it has been 
done in Ref.~\cite{alba-2015} for the Generalized Gibbs Ensemble (GGE) 
representative state.

For the local observables considered here, in each eigenstate expectation value 
$\langle\lambda|{\mathcal O}|\lambda\rangle$ in~\eref{qamc-obs} one can isolate 
the contribution of the different string sectors as 
%
\begin{equation}
\langle\lambda|{\mathcal O}|\lambda\rangle=\sum\limits_{n,\gamma}{\mathcal O}_n
(\lambda_{n;\gamma}), 
\end{equation}
%
where ${\mathcal O}_n$ is the contribution of the BGT $n$-strings to the expectation 
value of ${\mathcal O}$, and $\pm\lambda_{n;\gamma}$, with $\gamma$ labeling the 
different $n$-strings, are the solutions of the BGT equations~\eref{bgt-eq} 
identifying the Bethe state $|\lambda\rangle$. By comparing~\eref{qamc-obs} 
and~\eref{obs-th} one obtains that in the limit $L,N_{mcs}\to\infty$ 
%
\begin{equation}
\quad\fl\lim_{N_{mcs}\to\infty}\,\frac{1}{N_{mcs}}\sum\limits_{\lambda_{n;\gamma}}
{\mathcal O}_n(\lambda_{n;\gamma})\,\stackrel{L\to\infty}{\overrightarrow{\hspace{40pt}}}
\,\langle\pmb{\rho^*}|{\mathcal O}|\pmb{\rho^*}\rangle\equiv\sum_n\int_{-\infty}^{+\infty}d
\lambda\rho^*_{n}(\lambda){\mathcal O}_n(\lambda). 
\end{equation}
%
This suggests that the histogram of the $n$-strings BGT roots sampled in the Monte 
Carlo converges in the thermdynamic limit to the saddle point root distribution 
$\rho^*_n(\lambda)$.

This is demonstrated numerically in Figure~\ref{fig4:neel-rho} considering $\rho^*_1(\lambda)$ 
(panels (a)(b)) and $\rho^*_2(\lambda)$ (panel (c)(d)). The histograms correspond to Monte 
Carlo data for $L=48$ and $L=56$ sites. Panel (a) and (c) show the  histograms of the 
$1$-string and $2$-string BGT roots sampled in the Monte Carlo. The $y$-axis is rescaled 
by a factor $10^6$ for convenience. The width of the histogram bins $\Delta\lambda$ is 
$\Delta\lambda\approx0.02$ and $\Delta\lambda\approx0.001$ for $\rho_1(\lambda)$ and 
$\rho_2(\lambda)$, respectively. The histogram fluctuations are due both to the finite 
statistics (finite $N_{mcs}$) and to the finite size of the chain.

The extracted quench-action root distributions $\rho^*_1(\lambda)$ and $\rho^*(\lambda)$ 
are shown in panels (b) and (d). The data are the same as in panel (a)(c). The normalization 
of the distributions is chosen such as to match the analytical results from~\eref{rho1-sp} 
and~\eref{rho2-sp}, i.e., $\int d\lambda\rho^*_1(\lambda)\approx0.31$ and $\int d\lambda
\rho^*_2(\lambda)\approx0.015$. The Monte Carlo error bars shown in the Figure are obtained 
with a standard jackknife analysis~\cite{quenouille-1949,wolff-2004}. 
The continuous lines are the expected analytic result in the thermodynamic limit 
(cf.~\eref{rho1-sp}~\eref{rho2-sp}).

Clearly, the Monte Carlo data for $\rho^*_1(\lambda)$ are in excellent agreement 
with~\eref{rho1-sp} in the whole range $-2\le\lambda\le2$ considered. The statistical error 
bars are smaller than the symbol size. The oscillating (with $\lambda$) corrections 
visible at $|\lambda|\sim0.5$ are lattice effects, which decrease with the chain size 
(see the data for $L=48$). Much larger finite-size effects are observed for 
$\rho^*_2(\lambda)$ (panel (d) in the Figure). The corrections are larger on the tails of 
the root distribution. Moreover, the Monte Carlo error bars are clearly visible. This is 
due to the fact that since $\int d\lambda\rho^*_2(\lambda)/\sum_n\int d\lambda
\rho_n^*(\lambda)\approx 0.04$, the Monte Carlo statistics available to estimate 
$\rho_2^*(\lambda)$ is effectively reduced as compared to $\rho_1^*(\lambda)$. 
Finally, we numerically checked that these corrections are even larger for the $3$-strings,  
and much larger chains would be needed to extract $\rho^*_3(\lambda)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
\label{conclusions}

We presented a Monte Carlo implementation of the quench action method for integrable 
spin chains. We focused on the spin-$1/2$ isotropic Heisenberg ($XXX$) chain, considering  
the quench from the zero-momentum N\'eel state. The method is similar to the Monte Carlo 
approach developed in Ref.~\cite{alba-2015} to simulate the Generalized Gibbs Ensemble 
(GGE) in integrable models. The key idea is the Monte Carlo sampling of the chain Hilbert 
space with the quench action probability distribution given in~\eref{qa-prob}.

The approach heavily relies on the knowledge of the overlaps between the pre-quench N\'eel 
state $|N\rangle$ and the chain eigenstates, which have been obtained recently~\cite{
pozsgay-2014,brockmann-2014,brockmann-2014a,brockmann-2014b,brockmann-2014c,piroli-2014}. 
Moreover, the method is based on the detailed knowledge of the Hilbert space 
structure of the model provided by the Bethe ansatz formalism, in particular, on the 
so-called string hypothesis, and on the Bethe-Gaudin-Takahashi (BGT) equations. Although 
the approach is devised for finite-size systems, thermodynamic quantities can be extracted 
using finite-size scaling. Importantly, we restricted ourselves to a truncated Hilbert space, 
excluding the chain eigenstates containing zero-momentum strings. The reason is that 
zero-momentum strings lead to singularities in the N\'eel overlap formulas, which are 
tricky to deal with in the framework of the string hypothesis. Note that for the same 
reason the zero-momentum strings are neglected in the quench action approach~\cite{brockmann-2014} 
as well, although it has been argued that their effect should be irrelevant in the 
thermodynamic limit~\cite{brockmann-2014}.

In order to understand the effect of the zero-momentum strings in {\it finite} chains 
we first investigated the overlap distribution, including {\it all} the eigenstates 
with no zero-momentum strings, for chains up to $L=38$ sites. Interestingly, the total 
number of eigenstates having, in principle, non-zero N\'eel overlap is given in terms of 
the chain length by a simple combinatorial formula that we provided. Additional 
combinatorial formulas are provided for the case without zero-momentum strings, or 
at fixed particle number (i.e., magnetization), or fixed ``string content'' of the 
eigenstates. Surprisingly, for any finite chain the majority of eigenstates contain 
zero-momentum strings. Moreover, the fraction of eigenstates with nonzero-momentum 
strings vanishes as $L^{-1/2}$ in the thermdodynamic limit. This is reflected in the 
N\'eel overlap sum rules for the local conserved charges of the model. 
Although for any chain size their value is fixed by the N\'eel state expectation 
value, violations of the sum rules are observed, due to the Hilbert space truncation. 
Moreover, we numerically observed that the sum rules vanish as $L^{-1/2}$ in the 
thermodynamic limit. This behavior reflects the vanishing of the fraction of 
eigenstates with no zero-momentum strings, confirming that their contribution 
cannot be trivially neglected. This behavior as$L^{-1/2}$, however, is not generic, 
but it depends on the pre-quench initial state. This was demonstrated by 
considering the quench from the Majumdar-Ghosh state. We observed that the sum 
rules vanish as $1/L$, again reflecting the vanishing as $1/L$ of the fraction of 
eigenstates with no zero-momentum strings.

Next we presented a Monte Carlo implementation of the quench action method. 
We numerically demonstrated the validity of the approach, focusing on the N\'eel 
overlap sum rules for the local conserved quantities of the $XXX$ chain. Although 
for finite chains violations of the sum rules are present, their effect decays 
as $1/L$ upon increasing the chain size, meaning that overlap sum rules are 
restored in the thermodynamic limit. This implies that the only effect of the 
Hilbert space truncation is to give rise to scaling corrections. Physically, 
this means that the remaining eigenstates after the Hilber space truncation 
contain enough information about the post-quench thermodynamic behavior of the 
model. Finally, following Ref.~\cite{alba-2015} we extracted from the Monte Carlo 
the quench action BGT root distributions. Although finite-size oscillating 
corrections are present, already for relatively small chains with $L=56$ sites, 
the first two BGT root distributions are in impressive agreement with the 
analytic result in the thermodynamic limit.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Eigenstates with nonzero N\'eel overlap: eigenstates counting and 
string content}
\label{app-1}

Here we prove that the total number of eigenstates with N\'eel nozero overlap 
$Z_{Neel}(L)$for a chain of length $L$ is given as 
%
\begin{equation}
\label{N-count}
Z_{Neel}=2^{\frac{L}{2}-1}+\frac{1}{2}B\Big(\frac{L}{2},\frac{L}{4}\Big)+1. 
\end{equation}
%
For simplicity here we restrict ourselves to the situation with $L$ divisible by 
four. The strategy to prove~\eref{N-count} is to count all the possible parity-invariant 
BGT quantum numbers configurations. Let us consider the sector with fixed number of 
particles $M$, and a generic string content ${\mathcal S}=\{s_1,s_2,\dots,s_{M}\}$. 
Here $s_n$ is the number of $n$-strings, and one has the constraint $\sum_k ks_k=M$.

It is straightforward to check that total number of parity-invariant quantum number 
pairs ${\mathcal N}_n(L,{\mathcal S})$ in the $n$-string sector is given as 
%
\begin{equation}
\label{NnLS}
{\mathcal N}_n(L,{\mathcal S})=\Big\lfloor\frac{L}{2}-\frac{1}{2}
\sum_{m=1}^{M}t_{nm}s_m\Big\rfloor.
\end{equation}
%
where $t_{nm}\equiv 2\textrm{Min}(n,m)-\delta_{n,m}$. The number of parity-invariant 
quantum number configurations (i.e., eigenstates) ${\mathcal N}(L,{\mathcal S})$ 
compatible with string content ${\mathcal S}$ is obtained by choosing in all the 
possible ways the parity-invariant quantum number pairs independently in each 
$n$-string sector, which implies that    
%
\begin{equation}
\label{NLS}
{\mathcal N}(L,{\mathcal S})=\prod_{m=1}^{M} B\left({\mathcal N}_m,\left\lfloor
\frac{s_m}{2}\right\rfloor\right).
\end{equation}
%
Here the product is because each string sector is treated independently, while the 
factor $1/2$ in $s_m/2$ is because since all quantum numbers are organized in pairs, 
only half of the quantum numbers have to be specified. Note that in each $n$-string 
sector only one zero momentum (i.e., zero quantum number) string is allowed, due to 
the Pauli principle. Moreover, $s_m$ is odd (even) only if the zero momentum string 
is (not) present. The floor function $\lfloor\cdot\rfloor$ in~\eref{NLS} reflects 
that the quantum number of zero-momentum strings is fixed.

We now consider the string configurations with particle number $0\le\ell\le M$ and 
fixed number of strings $1\le q\le M/2$. Note that the maximul allowed string length 
is $M/2$ beacause of parity invariance. Note also that in determining $q$ strings of 
different length are treated equally. Clearly, one has that $\sum_m s_m=q$. 
For a given fixed pair $\ell,q$ the total number of quantum number configurations 
is given as 
%
\begin{equation}
\label{NLlq}
{\mathcal N}'(L,\ell,q)=\sum\limits_{\{\{s_m\}\,:\, \sum m s_m=\ell, \sum s_m=q\}}
{\mathcal N}(L,{\mathcal S}),
\end{equation}
%
where the sum is over the content $\{s_m\}_{m=1}^M$ compatible with the constraints 
$\sum_m s_m=q$ and $\sum_m m s_m=\ell$. The strategy is to write a recursive relation 
for ${\mathcal N}'(L,\ell,q)$. To this purpose it is useful to consider the shifted 
string content ${\mathcal S}'$ defined as  
%
\begin{equation}
{\mathcal S}'\equiv \{s_{m+1}\}\quad\textrm{with}\, s_m\in{\mathcal S},\,\forall m.
\end{equation}
%
Using the definition of $t_{ij}$, it is straightforward to derive that  
%
\begin{equation}
t_{ij}=t_{i-1,j-1}+2,
\end{equation}
%
which implies that ${\mathcal N}_n(L,{\mathcal S})$ (see~\eref{NnLS}) satisfies the 
recursive equation 
%
\begin{equation}
{\mathcal N}_n(L,{\mathcal S})={\mathcal N}_{n-1}(L-2q,{\mathcal S}'). 
\end{equation}
%
After substituting in~\eref{NLS} one obtains 
%
\begin{equation}
\label{NLSr}
{\mathcal N}(L,{\mathcal S})=B\Big({\mathcal N}_1(L,{\mathcal S}),\Big\lfloor \frac{s_1}{2}
\Big\rfloor\Big){\mathcal N}(L-2q,{\mathcal S}'). 
\end{equation}
%
Finally, after substituting~\eref{NLSr} in~\eref{NLlq}, one obtains a recursive relation 
for ${\mathcal N}'(L,\ell,q)$ as 
%
\begin{equation}
\label{NpLlq}
{\mathcal N}'(L,\ell,q)=\sum_{s=0}^{q-1}B\Big(\frac{L}{2}-q+\Big\lfloor\frac{s}{2}\Big\rfloor,
\left\lfloor\frac{s}{2}\right\rfloor\Big){\mathcal N}'\left(L-2q,\ell-q,
q-s\right), 
\end{equation}
%
with the constraint that when $\ell=q$ one has 
%
\begin{equation}
{\mathcal N}'(L,q,q)=B\Big(\Big\lfloor\frac{L-q}{2}\Big\rfloor,\Big\lfloor\frac{q}{2} 
\Big\rfloor\Big).
\end{equation}
%
This is obtained by observing that if $\ell=q$ only $1$-strings are allowed and~\eref{NnLS} 
gives ${\mathcal N}_n(L,{\mathcal S})=\lfloor (L-q)/2\rfloor$.

It is straightforward to check that for even $q$ the ansatz 
%
\begin{equation}
{\mathcal N}'(L,\ell,q)=\frac{q}{\ell}B\Big(\frac{L-\ell}{2},\frac{q}{2}\Big)
B\Big(\frac{\ell}{2},\frac{q}{2}\Big),
\end{equation}
% 
satisfies~\eref{NpLlq}. For odd $q$ the solution of~\eref{NpLlq} is 
%
\begin{equation}
{\mathcal N}'(L,\ell,q)=\frac{\ell-q+1}{\ell}B\Big(\frac{L-\ell}{2},\frac{q-1}{2}
\Big)B\Big(\frac{\ell}{2},\frac{q-1}{2}\Big).
\end{equation}
%
The number of eigenstates in the sector with $\ell$ particles with nonzero 
N\'eel overlap $Z'_{Neel}(L,\ell)$ are obtained by summing over all possible values 
of $q$ as 
%
\begin{equation}
\label{sum1}
Z'_{Neel}(L,\ell)=\sum\limits_{q=1}^\ell {\mathcal N}'(L,\ell,q).
\end{equation}
%
It is convenient to split the summation in~\eref{sum1} considering odd values of $q$ 
and even $q$ separately. For odd $q$ one obtains 
%
\begin{equation}
\sum\limits_{k=0}^{\ell/2-1} {\mathcal N}'(L,\ell,2k+1)=B\Big(\frac{L}{2}-1,
\frac{\ell}{2}-1\Big),
\end{equation}
%
while for even $q$ one has 
%
\begin{equation}
\sum\limits_{k=0}^{\ell/2} {\mathcal N}'(L,\ell,2k)=B\Big(\frac{L}{2}-1,
\frac{\ell}{2}\Big). 
\end{equation}
%
Putting everything together one obtains 
%
\begin{equation}
\label{N-count-p}
Z'_{Neel}(L,\ell)=B\Big(\frac{L}{2}-1,
\frac{\ell}{2}-1\Big)+B\Big(\frac{L}{2}-1,
\frac{\ell}{2}\Big). 
\end{equation}
%
The total number of eigenstates with nonzero N\'eel overlap $Z_{Neel}(L)$ 
(cf.~\eref{N-count}) is obtained from~\eref{N-count-p} by summing over 
the allowed values of $\ell=2k$ with $k=0,1,\dots,\ell/2$.

Note that the total number $Z_{MG}$ of parity-invariant eigenstates having non 
zero overlap with the Majumdar-Ghosh state is obtained from Eq~\eref{N-count-p} 
replacing $\ell=L/2$, to obtain 
%
\begin{equation}
\label{p-inv-mg}
Z_{MG}=B\Big(\frac{L}{2}-1,\frac{L}{4}-1\Big)+B\Big(\frac{L}{2}-1,\frac{L}{4}
\Big). 
\end{equation}
%
Physically, this is due to the fact that the 
Majumdar-Ghosh state is invariant under $SU(2)$ rotations, which implies 
that only eigenstates with zero total spin $S=0$ can have non zero overlap.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Excluding the zero-momentum strings}
\label{app-2}

Here we demonstrate that the total number of eigenstates with nonzero N\'eel overlap, 
which do not contain zero-momentum strings, $\widetilde Z_{Neel}(L)$ is given as 
%
\begin{equation}
\widetilde Z_{Neel}(L)=B\Big(\frac{L}{2},\frac{L}{4}\Big). 
\end{equation}
%
Given a generic $M$-particle eigenstate of the $XXX$ chain, due to parity invariance, 
if one excludes the zero-momentum strings only $n$-strings with length $n\le M/2$ 
are allowed. Similarly, the string content is of the form $\widetilde{\mathcal S}
\equiv\{\tilde s_1,\dots,\tilde s_{M/2}\}$, i.e., $\tilde s_m=0$ $\forall m>M/2$.
Note that due to parity invariance and to the exclusion of the zero-momentum strings 
one has that $\tilde s_m$ is always an even integer. Clearly one has $\sum_{m=1}^{M/2}
m \tilde s_m=M$.

The total number of parity-invariant quantum numbers $\widetilde{\mathcal N}_n$ in the 
$n$-string sector is given as  
%
\begin{equation}
\widetilde{\mathcal N}_n(L,\widetilde{\mathcal S})=\frac{L}{2}-\frac{1}{2}
\sum_{m=1}^{M/2}t_{nm}\tilde s_m.
\end{equation}
%
The proof now proceeds as in~\ref{app-1}. One can define the total number of eigenstates 
with nonzero N\'eel overla in the sector with $\ell$ particles and $q$ different strings as 
$\widetilde{\mathcal N}'(L,\ell,q)$. Note that due to parity invariance and the exclusion of 
zero-momentum strings, $q$ must be even. It is straigtforward to show that $\widetilde
{\mathcal N}'(L,\ell,q)$ obeys the recursive relation
%
\begin{equation}
\label{NpLlq-1}
\widetilde{\mathcal N}'(L,\ell,q)=\sum_{s=0}^{q/2-1}B\Big(\frac{L}{2}-q+s,s\Big)\widetilde
{\mathcal N}'\Big(L-2q,\frac{\ell-q}{2},\frac{q}{2}-s\Big),
\end{equation}
% 
with the constraint
%
\begin{equation}
\widetilde{\mathcal N}'(L,1,1)=\frac{L}{2}-1. 
\end{equation}
%
It is straightforward to check that the solution of~\eref{NpLlq-1} is given as 
%
\begin{equation}
\widetilde{\mathcal N}'(L,\ell,q)=\frac{L-2\ell+2}{L-\ell+2}B\Big(\frac{L-\ell}{2}+1,q\Big)
B\Big(\frac{\ell}{2}-1,\frac{q}{2}-1\Big).
\end{equation}
%
After summing over the allowed values of $q=2k$ with $k=1,2,\dots,\ell/2$ one obtains 
the total number of eigenstaets with nonzero N\'eel overlap at fixed number of 
particles $\ell$ $\widetilde Z_{Neel}'(L,\ell)$ as 
%
\begin{equation}
\label{neel-fi}
\widetilde Z_{Neel}'(L,\ell)=B\Big(\frac{L}{2},\frac{\ell}{2}\Big)-
B\Big(\frac{L}{2},\frac{\ell}{2}-1\Big).
\end{equation}
%
Summing over $\ell$ one obtains~\eref{neel-ov-count}.
Similar to~\eref{p-inv-mg} the total number of eigenstates $\widetilde Z_{MG}$ 
with no zero-momentum strings having non-zero overlap with the Majumdar-Ghosh 
state is obtained from~\eref{neel-fi} replacing $\ell\to L/2$, to obtain 
%
\begin{equation}
\label{mg-fi}
\widetilde Z_{MG}=B\Big(\frac{L}{2},\frac{L}{4}\Big)-B\Big(\frac{L}{2},\frac{L}{4}-1
\Big). 
\end{equation}
%
Interestingly, using~\eref{p-inv-mg} and~\eref{mg-fi}, one obtains that the ratio 
$\widetilde Z_{MG}/Z_{MG}$ is given as 
%
\begin{equation}
\frac{\widetilde Z_{MG}}{Z_{MG}}=\frac{4}{4+L}. 
\end{equation}
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Exact N\'eel and Majumdar-Ghosh overlaps for a small Heisenberg chain} 
\label{app-L12}

In this section we provide exact diagonalization results for the overlap of both the 
N\'eel state and the Majumdar-Ghosh (MG) state with all the eigenstates of the Heisenberg 
spin chain with $L=12$ sites. We also provide the corresponding results obtained 
using the string hypothesis and the overlap formulas~\eref{neel-ov} and~\eref{mg-ov}, 
restricting ourselves to eigenstates with no zero-momentum strings.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{N\'eel overlap}
\label{app-neel}

The overlaps between all the eigenstates of the Heisenberg spin chain and the N\'eel 
state are reported in Table~\ref{table:neel}. The first column in the Table shows 
the string content ${\mathcal S}\equiv\{s_1,\dots,s_M\}$, with $M$ being the number 
of finite rapidities. The number of infinite rapidities $N_{\infty}=L/2-M$ is also 
reported. Note that eigenstates containing infinite rapidities correspond to 
different $S_z$ eigenvalue. The second column shows 
$2I_n^+$, with $I_n$ the Bethe-Gaudin-Takahashi quantum number identifying the 
BGT rapidity of the $n$-string. Due to the parity invariance only the positive 
quantum numbers are reported. The total number of independent strings, i.e., 
$q\equiv\sum_js_j$, is reported in the third column. The fourth column is the 
eigenstate's energy eigenvalue $E$. The last two columns show the squared N\'eel 
overlaps and the corresponding result obtained using the Bethe-Gaudin-Takahashi 
equations, respectively. In the last column only the case with no zero-momentum 
strings is considered. The deviations from the exact diagonalization results (digits 
with different colors) have to be attributed to the string hypothesis. Notice that  
the overlap between the N\'eel state and the $S_z=0$ eigenstate in the sector with 
maximal total spin $S=L/2$ (first column in Table~\ref{table:neel}), is given 
analytically as $2/B(L,L/2)$, with $B(x,y)$ the Newton binomial.

Figure~\ref{fig1-BGT-check} plots the squared overlaps $|\langle\lambda|
N\rangle|^2$ between the N\'eel state and the eigenstates of the chain. 
The overlaps are plotted against the eigenstate energy density $E/L
\in[-\log(2),0]$. The circles are exact diagonalization results for all 
the chain eigenstates ($382$ eigenstates), whereas the crosses denote the 
overlaps calculated using formula~\eref{neel-ov}, and the 
Bethe-Gaudin-Takahashi equations. Note that only the eigenstates with 
no zero-momentum strings are shown ($252$ eigenstates) in the Figure. 
Panel (a) in the Figure is an overview of all the results. Panels (b)-(d) 
correspond to zooming to the smaller overlap values $|\langle N|\lambda
\rangle|\lesssim 0.02$, $|\langle N|\lambda\rangle|\lesssim 0.002$, and 
$|\langle N|\lambda\rangle|\lesssim 10^{-5}$.

Clearly, the overlaps decay rapidly upon increasing the energy density. 
This is expected since the $XXX$ Hamiltonian expectation value over 
the N\'eel state is $\langle N|{H}|N\rangle=-1/2$. Importantly, the 
agreement between the exact diagonalization results and the results 
obtained using the BGT equations~\eref{ba_eq} is excellent, confirming 
the validity of the string hypothesis.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
\scriptsize
\centering
Bethe states with nonzero N\'eel overlap ($L=12$)\\[1ex]
\begin{tabular}{rrrrrr}
\toprule
String content & $2I^+_n$ & $q$ & E & $|\langle\lambda|N\rangle|^2$ (exact) & $|\langle
\lambda|N\rangle|^2$ (BGT) \\[0.3em]
\toprule
6 inf & - & - & $0$ & $0.002164502165$ & $0.002164502165$\\
\midrule
\{2,0\}\, 4 inf &$1_1$ & $2$ & $-3.918985947229$ & $0.096183409244$ & $0.096183409244$\\
 &$3_1 $ & & $-3.309721467891$ & $0.011288497947$ &             $0.011288497947$\\
 &$5_1 $ & & $-2.284629676547$ & $0.004542580506$ &             $0.004542580506$\\
 &$7_1 $ & & $-1.169169973996$ & $0.002752622983$ &             $0.002752622983$\\
 &$9_1 $ & & $-0.317492934338$ & $0.002116006203$ &             $0.002116006203$\\
\midrule
\{4,0,0,0\}\, 2 inf &$1_1 3_1 $ & $4$ & $-7.070529325964$ & $0.310133033838$ &$0.310133033838$\\
  &$1_1 5_1 $ & & $-5.847128730477$ & $0.129277023687$ &           $0.129277023687$\\
  &$ 1_1 7_1$ & & $-4.570746557876$ & $0.085992436024$ &           $0.085992436024$\\
  &$ 3_1 5_1$ & & $-5.153853093221$ & $0.015256395523$ &           $0.015256395523$\\
  &$3_1 7_1 $ & & $-3.916336243695$ & $0.010091113504$ &           $0.010091113504$\\
  &$5_1 7_1 $ & & $-2.817696043731$ & $0.004059780228$ &           $0.004059780228$\\
\midrule
\{0,2,0,0\}\, 2 inf &$1_2 $ & $2$ & $-1.905667167442$ & $0.001207238321$ & $0.0012072{\color{red}45406}$\\
  &$3_2 $ & & $-1.368837200825$ & $0.002340453815$ &            $0.0023{\color{red}25724713}$\\
  &$5_2 $ & & $-0.681173793635$ & $0.001921010489$ &            $0.0019{\color{red}39001396}$\\
\midrule
\{1,0,1,0\}\, 2 inf &$0_1 0_3$ & $2$ & $-2.668031843135$ & $0.034959609810$ & -\\
\midrule
\{6,0,0,0,0,0\}\, 0 inf &$1_1 3_1 5_1$ & $6$ & $-8.387390917445$ & $0.153412152966$ & $0.153412152966$\\
\midrule
\{2,2,0,0,0,0\}\, 0 inf &$1_1 1_2$ & $4$ & $-5.401838225870$ & $0.040162686361$ & $0.04{\color{red}1042488913}$\\  
&$3_1 1_2 $ & & $-4.613929948329$ & $0.004636541934$ & $0.004{\color{red}730512604}$\\
  &$5_1 1_2 $ &  & $-3.147465758841$ & $0.001335522556$ & $0.00133{\color{red}7334035}$\\
\midrule
\{3,0,1,0,0,0\}\, 0 inf &$0_1 2_1 0_3$ & $4$ & $-6.340207488736$ & $0.052743525774$ & -\\
  &$0_1 4_1 0_3$ & & $-5.203653009936$ & $0.015022005621$ & - \\
  &$0_1 6_1 0_3$ & & $-3.788693957250$ & $0.011144489334$ & - \\
\midrule
\{1,0,0,0,1,0\}\, 0 inf &$0_1 0_5$ & $2$ & $-2.444293750583$ & $0.005887902992$ & - \\
\midrule
\{0,0,2,0,0,0\}\, 0 inf &$1_3$ & $2$ & $-1.111855930538$ & $0.001342476001$ & $0.0013{\color{red}84980817}$ \\
\midrule
\{0,1,0,1,0,0\}\, 0 inf &$0_2 0_4$ & $2$ &  $-1.560671012472$ & $0.000026982174$ & - \\
\bottomrule
\end{tabular}
\caption{All Bethe states for $L=12$ having nonzero overlap with the zero-momentum N\'eel state. 
 The first column shows the string content of the Bethe states, including the number of infinite 
 rapidities. The second and third column show $2I_n^+$, with $I_n^+$ the BGT quantum numbers 
 identifying the different states, and the number $q$ of independent strings. In the second 
 column only the positive BGT numbers are shown. The fourth column is the Bethe state eigenenergy. 
 Finally, the last two columns show the exact overlap with the N\'eel state and the approximate 
 result obatained using the BGT equations. In the last column Bethe states containing zero-momentum 
 strings are excluded. Deviations from the exact result (digits with different colors) are 
 attributed to the string hypothesis. 
}
\label{table:neel}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%##################################################################
\begin{figure}[t]
\begin{center}
\includegraphics[width=.9\textwidth]{./draft_figs/L20_BT_check}
\end{center}
\caption{ The squared overlap $|\langle N|\lambda\rangle|^2$ between the the 
 Neel state $|N\rangle$ and the eigenstates $|\lambda\rangle$ of the $XXX$ 
 chain with $L=20$ sites. Only non-zero overlaps are shown. In all the panels the 
 $x$-axis shows the eigenstate energy density $E/L$. The circles are the exact 
 diagonalization results for all the non-zero overlaps. The crosses are the Bethe 
 ansatz results obtained using the Bethe-Gaudin-Takahashi equations. The missing 
 crosses correspond to eigenstates containing zero-momentum strings. (a) Overview 
 of all the non-zero overlaps. (b)(c)(d) The same overlaps as in (a) zooming in 
 the regions $[0,0.2]$, $[0,0.020]$, and $[0,4\cdot 10^{-5}]$. The discrepancies 
 between the ED and the Bethe ansatz results are attributed to the string 
 deviations. 
}
\label{fig1-BGT-check}
\end{figure}
%##################################################################

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Majumdar-Ghosh overlap}
\label{app-mg}
The overlap between the Heisenberg chain eigenstates with the Majumdar-Ghosh state are shown in 
Table~\ref{table:mg}. The conventions on the representation of the eigenstates is the same as in 
Table~\ref{table:neel}. Note that in contrast with the N\'eel state, only the eigenstates with 
zero total spin $S=0$ have non zero overlap, i.e., no eigenstates with infinite rapidities are 
present, which reflect that the Majumdar-Ghosh state is unvariant under $SU(2)$ rotations.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
\scriptsize
\centering
Bethe states with nonzero N\'eel overlap ($L=12$)\\[1ex]
\begin{tabular}{rrrrrr}
\toprule
String content & $2I^+_n$ & $q$ & E & $|\langle\lambda|MG\rangle|^2$ (exact) & $|\langle\lambda|MG\rangle|^2$ (BGT) \\[0.3em]
\toprule
\{6,0,0,0,0,0\} &$1_1 3_1 5_1$ & $6$ & $-8.387390917445$ & $0.716615769224$ & $0.716615769224$\\
\midrule
\{2,2,0,0,0,0\} &$1_1 1_2$ & $4$ & $-5.401838225870$ & $0.055624700196$ & $0.05{\color{red}4033366543}$\\  
&$3_1 1_2 $ & & $-4.613929948329$ & $0.005687428810$ & $0.005{\color{red}582983043}$\\
&$5_1 1_2 $ &  & $-3.147465758841$ & $0.002107475934$ & $0.002107{\color{red}086933}$\\
\midrule
\{3,0,1,0,0,0\} &$0_1 2_1 0_3$ & $4$ & $-6.340207488736$ & $0.205891158647$ & -\\
  &$0_1 4_1 0_3$ & & $-5.203653009936$ & $0.038832154450$ & - \\
  &$0_1 6_1 0_3$ & & $-3.788693957250$ & $0.006019410923$ & - \\
\midrule
\{1,0,0,0,1,0\} &$0_1 0_5$ & $2$ & $-2.444293750583$ & $0.000129601311$ & - \\
\midrule
\{0,0,2,0,0,0\} &$1_3$ & $2$ & $-1.111855930538$ & $0.000011727787$ & $0.00001{\color{red}2785580}$\\
\midrule
\{0,1,0,1,0,0\} &$0_2 0_4$ & $2$ &  $-1.560671012472$ & $0.000330572718$ & - \\
\bottomrule
\end{tabular}
\caption{All Bethe states for $L=12$ having nonzero overlap with the zero-momentum Majumdar-Ghosh (MG) 
 state. The first column shows the string content of the Bethe states. The second and third column show 
 $2I_n^+$, with $I_n^+$ the BGT quantum numbers identifying the different states, and the number $q$ 
 of independent strings. In the second column only the positive BGT numbers are shown. Note that, in 
 contrast to Table~\ref{table:neel} no states with infinite rapidities are present. The fourth column 
 is the Bethe state eigenenergy. Finally, the last two columns show the exact overlap with the MG state 
 and the approximate result obatained using the BGT equations. In the last column Bethe states containing 
 zero-momentum strings are excluded. Deviations from the exact result (digits with different colors) 
 are attributed to the string hypothesis. 
}
\label{table:mg}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{thebibliography}{99}

\bibitem{bloch-2008}
I.~Bloch, J.~Dalibard, and W.~Zwerger, Rev.\ Mod.\ Phys.\ {\bf 80},
885 (2008).

\bibitem{greiner-2002}
M.~Greiner, O.~Mandel, T. H\"ansch, and I.~Bloch, Nature (London)
{\bf 419}, 51 (2002).

\bibitem{kinoshita-2006}
T.~Kinoshita, T.~Wenger, and D.~S.~Weiss, Nature (London) {\bf 440},
900 (2008).

\bibitem{hofferberth-2007}
S.~Hofferberth, I.~Lesanovsky, B.~Fischer, T.~Schumm, and J.~Schiedmayer,
Nature (London) {\bf 449}, 324 (2007).

\bibitem{trotzky-2012}
S.~Trotzky, Y.-A.~Chen, A.~Flesch, I.~P.~McCulloch, U.~Schollw\"ock,
J.~Eisert, and I.~Bloch, Nature Phys.\ {\bf 8}, 325 (2012).

\bibitem{gring-2012}
M.~Gring, M.~Kuhnert, T.~Langen, T.~Kitagawa, B.~Rauer, M.~Schreitl,
I.~Mazets, D.~A.~Smith, E.~Demler, and J.~Schmiedmayer, Science {\bf 337},
6100 (2012).

\bibitem{cheneau-2012}
M.~Cheneau, P.~Barmettler, D.~Poletti, M.~Endres, P.~Schaua, T.~Fukuhara,
C.~Gross, I.~Bloch, C.~Kollath, and S.~Kuhr, Nature (London) {\bf 481},
484 (2012).

\bibitem{schneider-2012}
U.~Schneider, L.~Hackeruller, J.~P.~Ronzheimer, S.~Will, S.~Braun, T.~Best,
I.~Bloch, E.~Demler, S.~Mandt, D.~Rasch, and A.~Rosch, Nature\ Phys.\
{\bf 8}, 213 (2012).

\bibitem{kunhert-2013}
M.~Kuhnert, R.~Geiger, T.~Langen, M.~Gring, B.~Rauer,
T.~Kitagawa, E.~Demler, D.~Adu Smith, and J.~Schmiedmayer, Phys.\ Rev.\
Lett.\ {\bf 110}, 090405 (2013).

\bibitem{langen-2013}
T.~Langen, R.~Geiger, M.~Kuhnert, B.~Rauer, and J.~Schmiedmayer,
Nature\ Phys.\ {\bf 9}, 640 (2013).

\bibitem{meinert-2013}
F.~Meinert, M.~J.~Mark, E.~Kirilov, K.~Lauber, P.~Weinmann,
A.~J.~Daley, and H.-C.~Nagerl, Phys.\ Rev.\ Lett.\ {\bf 111},
053003 (2013).

\bibitem{fukuhara-2013}
T.~Fukuhara, A.~Kantian, M.~Endres, M.~Cheneau, P.~Schaua, S.~Hild, C.~Gross,
U.~Schollw\"ock, T.~Giamarchi, I.~Bloch, and S.~Kuhr, Nature\ Phys.\ {\bf 9},
235 (2013).

\bibitem{ronzheimer-2013}
J.~P.~Ronzheimer, M.~Schreiber, S.~Braun, S.~S.~Hodgman, S.~Langer, I.~P.~McCulloch,
F. Heidrich-Meisner, I.~Bloch, and U.~Schneider, Phys.\ Rev.\ Lett.\ {\bf 110},
205301 (2013).

\bibitem{braun-2014}
S.~Braun, M.~Friesdorf, S.~Hodgman, M.~Schreiber, J.~Ronzheimer, A.~Riera, M.~del Rey,
I.~Bloch, J.~Eisert, and U.~Schneider, PNAS {\bf 112}, 3641 (2015).

\bibitem{langen-2015}
T.~Langen, S.~Erne, R.~Geiger, B.~Rauer, T.~Schweigier, M.~Kuhnert, W.~Rohringer,
I.~E.~Mazets, T.~Gasenzer, J.~Schmiedmayer, Science {\bf 348}, 6231 (2015).

\bibitem{polkovnikov-2011}
A.~Polkovnikov, K.~Sengupta, A~Silva, and M.~Vengalattore, Rev.\ Mod.\ Phys.\
{\bf 83}, 863 (2011).

\begin{comment}
\bibitem{calabrese-2006}
P.~Calabrese and J.~Cardy, Phys.\ Rev.\ Lett.\ {\bf 96}, 136801 (2006).

\bibitem{rigol-2007}
M.~Rigol, V.~Dunjko, V.~Yurovsky, and M.~Olshanii, Phys.\ Rev.\ Lett.\ 
{\bf 98}, 050405 (2007).

\bibitem{calabrese-2007}
P.~Calabrese and J.~Cardy, J.\ Stat.\ Mech.\ (2007) P06008.

\bibitem{kollath-2007}
C.~Kollath, A.~M.~L\"auchli, and E.~Altman, Phys.\ Rev.\ Lett.\ 
{\bf 98}, 180601 (2007).

\bibitem{manmana-2007}
S.~R.~Manmana, S.~Wessel, R.~M.~Noack, and A.~Muramatsu, 
Phys.\ Rev.\ Lett.\ {\bf 98}, 210405 (2007).

\bibitem{rigol-2008}
M.~Rigol, V.~Dunjko, and M.~Olshanii, Nature {\bf 452}, 854 (2008).

\bibitem{cramer-2008}
M.~Cramer, C.~M.~Dawson, J.~Eisert, and T.~J.~Osborne, Phys.\ Rev.\ 
Lett.\ {\bf 100}, 030602 (2008).

\bibitem{barthel-2008}
T.~Barthel and U.~Schollw\"ock, Phys.\ Rev.\ Lett.\ {\bf 100}, 100601 
(2008).

\bibitem{kollar-2008}
M.~Kollar and M.~Eckstein, Phys.\ Rev.\ A {\bf 78}, 013626 (2008).

\bibitem{moeckel-2008}
M.~Moeckel and S.~Kehrein, Phys.\ Rev.\ Lett.\ {\bf 100}, 175702 (2008).

\bibitem{iucci-2009}
A.~Iucci and M.~A.~Cazalilla, Phys.\ Rev.\ A\ {\bf 80}, 063619 (2009).

\bibitem{rossini-2009}
D.~Rossini, A.~Silva, G.~Mussardo, and G.~E.~Santoro, Phys.\ Rev.\ 
Lett.\ {\bf 102}, 127204 (2009).

\bibitem{barmettler-2009}
P.~Barmettler, M.~Punk, V.~Gritsev, E.~Demler, and E.~Altman, Phys.\ Rev.\ 
Lett.\ {\bf 102}, 130603 (2009).

\bibitem{biroli-2010}
G.~Biroli, C.~Kollath, and A.~M.~L\"auchli, Phys.\ Rev.\ Lett.\ 
{\bf 105}, 250401 (2010).

\bibitem{rossini-2010}
D.~Rossini, S.~Suzuki, G.~Mussardo, G.~E.~Santoro, and A.~Silva, 
Phys.\ Rev.\ B\ {\bf 82}, 144302 (2010).

\bibitem{fioretto-2010}
D.~Fioretto and G.~Mussardo, New\ J.\ Phys.\ {\bf 12}, 
055015 (2010).

\bibitem{gogolin-2011}
C.~Gogolin, M.~P.~Mueller, and J.~Eisert, Phys.\ Rev.\ Lett.\ 
{\bf 106}, 040401 (2011).

\bibitem{banuls-2011}
M.~C.~Ba\~nuls, J.~I.~Cirac, and M.~B.~Hastings, Phys.\ Rev.\ Lett.\ 
{\bf 106}, 050405 (2011).

\bibitem{calabrese-2011}
P.~Calabrese, F.~H.~L.~Essler, and M.~Fagotti, Phys.\ Rev.\ Lett.\ {\bf106}, 227203 (2011).

\bibitem{rigol-2011}
M.~Rigol and M.~Fitzpatrick, Phys.\ Rev.\ A {\bf 84}, 033640 (2011).

\bibitem{calabrese-2012}
P.~Calabrese, F.~H.~L.~Essler, and M.~Fagotti, J.\ Stat.\ Mech.\ (2012) P07016.

\bibitem{caux-2012}
J.-S.~Caux and R.~M.~Konik, Phys.\ Rev.\ Lett.\ {\bf 109}, 175301 (2012).

\bibitem{essler-2012}
F.~H.~L.~Essler, S.~Evangelisti, and M.~Fagotti, Phys.\ Rev.\ Lett.\ 
{\bf 109}, 247206 (2012).

\bibitem{cazalilla-2012}
M.~A.~Cazalilla, A.~Iucci, and M.-C.~Chung, Phys.\ Rev.\ E {\bf 85}, 
011133 (2012).

\bibitem{mossel-2012a}
J.~Mossel and J.-S.~Caux, New\ J.\ Phys.\ {\bf 14} 075006 (2012).

\bibitem{collura-2013}
M.~Collura, S.~Sotiriadis and P.~Calabrese, Phys.\ Rev.\ Lett.\ {\bf 110}, 245301 (2013)

\bibitem{mussardo-2013}
G.~Mussardo, Phys.\ Rev.\ Lett.\ {\bf 111}, 100401 (2013).

\bibitem{caux-2013}
J.-S.~Caux and F.~H.~L.~Essler, Phys.\ Rev.\ Lett.\ {\bf 110}, 
257203 (2013).

\bibitem{fagotti-2013a}
M.~Fagotti and F.~H.~L.~Essler, J.\ Stat.\ Mech.\ (2013), P07012.

\bibitem{fagotti-2013}
M.~Fagotti and F.~H.~L.~Essler, Phys.\ Rev.\ B\ {\bf87}, 245107 (2013).

\bibitem{sotiriadis-2014}
S.~Sotiriadis and P.~Calabrese, J.\ Stat.\ Mech.\ (2014) P07024.

\bibitem{fagotti-2014}
M.~Fagotti, M.~Collura, F.~H.~L.~Essler, and P.~Calabrese, Phys.\ Rev.\ B {\bf 89}, 
125101 (2014).

\bibitem{essler-2014}
F.~H.~L.~Essler, S.~Kehrein, S.~R.~Manmana, and N.~J.~Robinson, Phys.\ Rev.\ B {\bf 89}, 
165104 (2014).

\bibitem{goldstein-2014}
G.~Goldstein and N.~Andrei, arXiv:1405.4224.

\bibitem{de-nardis-2014}
J.~De Nardis, B.~Wouters, M.~Brockmann, and J.-S.~Caux, Phys.\ Rev.\ A {\bf 89}, 
033601 (2014).

\bibitem{wright-2014}
T.~M.~Wright, M.~Rigol, M.~J.~Davis, and K.~V.~Kheruntsyan, Phys.\ Rev.\ Lett.\ {\bf 113}, 
050601 (2014).

\bibitem{pozsgay-2014}
B.~Pozsgay, M.~Mesty\'an, M.~A.~Werner, M.~Kormos, G.~Zar\'and, and G.~Tak\'acs,
Phys.\ Rev.\ Lett.\ {\bf 113}, 117203 (2014).

\bibitem{wouters-2014}
B.~Wouters, M.~Brockmann, J.~De~Nardis, D.~Fioretto, M.~Rigol, and J.-S.~Caux, 
Phys.\ Rev.\ Lett.\ {\bf 113}, 117202 (2014).

\bibitem{mestyian-2015}
M~Mesty\'an, B.~Pozsgay, G.~Tak\'acs, and M.~A.~Werner, J.\ Stat.\ Mech.\ (2015) 
P04001.

\bibitem{ilievski-2015a}
E.~Ilieveski, J.~De~Nardis, B.~Wouters, J.-S.~Caux, F.~H.~Essler, and T.~Prosen, 
arXiv:1507.02993.

\bibitem{kcc14}
M.~Kormos, M.~Collura, and P.~Calabrese, Phys.\ Rev.\ A {\bf 89}, 013609 
(2014).

\bibitem{kcc14a}
P.~P.~Mazza, M.\ Collura, M.\ Kormos, and P.\ Calabrese, J.\ Stat.\ Mech.\  
(2014) P11016.

\bibitem{prosen-2014}
T.~Prosen, Nucl.\ Phys.\ B\ {\bf 886}, (2014) 1177.

\bibitem{pereira-2014}
R.~G.~Pereira, V.~Pasquier, J.~Sirker, and I.~Affleck, J.\ Stat.\ Mech.\ 
(2014) P09037.

\bibitem{ilievski-2015}
E.~Ilievski, M.~Medejak, and T.~Prosen, arXiv:1506.05049.

\bibitem{essler-2015}
F.~H.~L.~Essler, G.~Mussardo, and M.~Panfil, Phys.\ Rev.\ A {\bf 91}, 051602 (2015).

\bibitem{cardy-2015}
J.~Cardy,  arXiv:1507.07266.

\bibitem{iyer-2015}
D.~Iyer, M.~Srednicki, and M.~Rigol, Phys.\ Rev.\ E\ {\bf 91}, 062142 (2015).

\bibitem{takahashi-1971} 
M.~Takahashi, Prog.\ Theor.\ Phys.\ {\bf 46}, 401 (1971).

\end{comment}

\bibitem{taka-book}
M.~Takahashi, {\it Thermodynamics of one-dimensional solvable models}, 
Cambridge University Press, Cambridge, 1999.

\bibitem{gu-2005}
S.-J.~Gu, N.~M.~R.~Peres, Y.-Q.~Li, Eur.\ Phys.\ J.\ B {\bf 48}, 157 (2005).

\bibitem{grabowski-1995}
M.~P.~Grabowski and P.~Mathieu, Ann.\ Phys.\ N.Y. {\bf 243}, 
299 (1995).

\bibitem{bethe-1931}
H.~Bethe, Z.\ Phys.\ {\bf 71}, 205 (1931).

\bibitem{mossel-2012}
J.~Mossel and J.-S.~Caux, J.\ Phys.\ A:\ Math.\ Theor.\ {\bf 45}, 
255001 (2012).

\bibitem{pozsgay-2013}
B.~Pozsgay, J.\ Stat.\ Mech.\ (2013) P07003.

\bibitem{kor-book}
V.~E.~Korepin, N.~M.~Bogoliubov, and A.~G.~Izergin, \emph{Quantum 
Inverse Scattering Methods and Correlation Functions}, Cambridge 
University Press, Cambridge, 1997.

\bibitem{landau-binder}
D.~P.~Landau and K.~Binder, \emph{A Guide to Monte Carlo Simulations in 
Statistical Physics}, Cambridge University Press, Cambridge, 2000.

\bibitem{faddeev-1996}
L.~D.~Faddeev, arXiv:9605187.

\bibitem{yang-1969}
C.~N.~Yang and C.~P.~Yang, J.\ Math.\ Phys. {\bf 10}, 1115 (1969).

\bibitem{calabrese-2014}
P.~Calabrese and P.~Le Doussal, J.\ Stat.\ Mech.\ (2014) P05004.

\bibitem{alba-2015}
V.~Alba, arXiv:1507.06994.

\bibitem{bethe-1931}
H.~Bethe, Zur Theorie der Metalle. I. Eigenwerte und Eigenfunktionen 
der linearen Atomkette, Z.\ Phys.\ {\bf 71}, 205 (1931).

\bibitem{brockmann-2014}
M.~Brockmann, J.~De~Nardis, B.~Wouters, and J.-S.~Caux, J.\ Phys.\ A:\ 
Math.\ Theor. {\bf 47}, 345003 (2014).

\bibitem{brockmann-2014c}
M.~Brockmann, J.~De~Nardis, B.~Wouters, and J.-S.~Caux, J.\ Phys.\ A:\ 
Math.\ Theor. {\bf 47}, 145003 (2014).

\bibitem{brockmann-2014b}
M.~Brockmann, J.\ Stat.\ Mech.\ (2014) P05006.

\bibitem{pozsgay-2014}
B.~Pozsgay, J.\ Stat.\ Mech.\ (2014) P06011.

\bibitem{calabrese-2007}
P.~Calabrese and J.-S.~Caux, Phys.\ Rev.\ Lett.\ {\bf 98}, 150403 (2007).

\bibitem{calabrese-2007-a}
P.~Calabrese and J.-S.~Caux, J.\ Stat.\ Mech.\ P08032 (2007).

\bibitem{de-nardis-2014}
J.~De Nardis, B.~Wouters, M.~Brockmann, and J.-S.~Caux, Phys.\ Rev.\ A {\bf 89}, 
033601 (2014).

\bibitem{pozsgay-2014}
B.~Pozsgay, M.~Mesty\'an, M.~A.~Werner, M.~Kormos, G.~Zar\'and, and G.~Tak\'acs,
Phys.\ Rev.\ Lett.\ {\bf 113}, 117203 (2014).

\bibitem{wouters-2014}
B.~Wouters, M.~Brockmann, J.~De~Nardis, D.~Fioretto, M.~Rigol, and J.-S.~Caux, 
Phys.\ Rev.\ Lett.\ {\bf 113}, 117202 (2014).

\bibitem{mestyian-2015}
M~Mesty\'an, B.~Pozsgay, G.~Tak\'acs, and M.~A.~Werner, J.\ Stat.\ Mech.\ (2015) 
P04001.

\bibitem{ilievski-2015a}
E.~Ilieveski, J.~De~Nardis, B.~Wouters, J.-S.~Caux, F.~H.~Essler, and T.~Prosen, 
arXiv:1507.02993.

\bibitem{brockmann-2014a}
M.~Brockmann, B.~Wouters, D.~Fioretto, J.~De Nardis, R.~Vlijm, and J.-S.~Caux, 
J.\ Stat.\ Mech.\ (2014) P12009.

\bibitem{mazza-2015}
P.~P.~Mazza, J.-M.~St\'ephan, E.~Canovi, V.~Alba, M.~Brockmann, and M.~Haque, 
arXiv:1509.04666.

\bibitem{fagotti-2013}
M.~Fagotti and F.~H.~Essler, J.\ Stat.\ Mech.\ (2013) P07012.

\bibitem{grabowski-1995}
M.~P.~Grabowski and P.~Mathieu, Ann.\ Phys.\ N.Y. {\bf 243},
299 (1995).

\bibitem{quenouille-1949} 
M.~H.~Quenouille, Ann.\ Math.\ Statist.\ {\bf 20}, 355 (1949).

\bibitem{wolff-2004} 
U.~Wolff, Comput.\ Phys.\ Comm.\ {\bf 156}, 143 (2004).

\bibitem{piroli-2014}
L.~Piroli and P.~Calabrese, J.\ Phys.\ A {\bf 47}, 385003 (2014).

\end{thebibliography}

\end{document}