### 2016-01-25 QAMC_draft/referee_response/response.txt
response

Dear editor,

We are submitting a revised version of our manuscript "The quench action approach 
in finite integrable spin chains" in which we address all the points raised 
by the referee.

Sincerely,

Vincenzo Alba
Pasquale Calabrese

---------------------------------------------------------------------
RESPONSE TO THE REFEREE
---------------------------------------------------------------------

REFEREE:

This paper represents a numerical implementation of the quench action approach,
introduced in [Phys. Rev. Lett. 110, 257203 (2013)], and in particular
addresses some questions raised in [J. Stat. Mech. (2014) P05004, Phys.
Rev. Lett. 113, 117202 (2014), J. Stat. Mech. (2014) P12009], concerning
the applicability of the method to interacting integrable models (where the
scattering matrix allows for the creation of bound states). These questions
are answered in full details and the developed numerical algorithm can be
generalized to other studies of the XXZ model in- and out-of-equilibrium.

RESPONSE:

	We thank the referee for the appreciating our work and for the positive 
	evaluation.

REFEREE:

	Abstract: There is probably a typo: ”charged” should indeed be ”charges”.

RESPONSE:

	We corrected the typo in the new version of the manuscript.

REFEREE:

	After formula 2: mutually conserved charges may be replaced with
	”mutually conserved local (and quasi-local) charges” which are the only
	one relevant for the application of the GGE logic.

RESPONSE:

	We modified the sentence according to the referee suggestion.

REFEREE:

	Page 4 : ” Physically, zero-momentum strings correspond to eigenstates
	amplitudes containing multi-particle bound states which are fully delocalized
	in the chain”. All the strings in each eigenstate are fully delocalized
	being eigenstate of the momentum operator. One may say that
	the zero momentum ones correspond to bound states of spins with zero
	propagation velocity (the post-quench light-cone velocity associated to
	them is zero).

RESPONSE:

	We thank the referee for the useful suggestion. We rephrased the sentence 
	in the new version of the manuscript to meet the referee suggestion.

REFEREE:

	Page 6 : ”the string deviations, which typically, i.e., for most of the
	chain eigenstates, vanish exponentially with L in the thermodynamic
	limit”. One may add that the complications arise when this condition
	is not valid, namely for zero momentum strings, when the deviation
	decay as power laws.

RESPONSE:

	We would like to thank the referee for the useful suggestion. We now 
	mention in the manuscript that zero-momentum strings exhibit deviations 
	which decay with a power law.

REFEREE:

	Page 8 : the observable ${\mathcal O}$ in formula 20 must be local for its matrix
	elements to be described by only a set of macroscopic distributions.

RESPONSE:

	We now rephrased the sentence as

		Moreover, for all the local observables ${\mathcal O}$ (the ones considered here)
		the contribution of the ...

	to stress the importance of the locality or quasilocality of ${\mathcal O}$

REFEREE:

	Formula 21: please define B (it is indeed defined later but here
	appears for the first time).

RESPONSE:

	We now define the Newton binomial after formula 21.

REFEREE:

	Formula 57: better use the same notation as in the determinant for the 
	string centers ($\lambda_{n;j}) since j is already used for the complex 
	rapidities)

RESPONSE:

	We modified the notation as suggested by the referee in formula 57 and 
	formula 59 as well.

REFEREE:

	After formula 71: One may comment that the vanishing of the overlaps
	with zero momentum even strings is related to the conservation laws
	of quasi-local charges [Phys. Rev. Lett. 115, 157201 (2015)]

RESPONSE:

	We now added the sentence after formula 71 as

		"This is because parity-invariant eigenstates with a single zero-momentum even-length ^M
		string, which are included in~\eref{zNeel1}, have identically zero N\'eel ^M
		overlap~\cite{brockmann-2014}. This is not related to the symmetries of the N\'eel ^M
		state, but to an ``accidental'' vanishing of the prefactor in the overlap ^M
		formula~\eref{Neel-ov}. In particular, this is related to the conservation of quasi-local
		charges~\cite{ilievski-2015a}."

	to mention the relationship between vanishing of the overlaps with zero-momentum strings 
	and conservation of quasi-local charges.

REFEREE:

	section 5.2 contains repetitions and numerous overlaps with section 2.4,
	maybe a rearrangement of the content of the two sections is necessary?

RESPONSE:

	We understand the point of the referee. However, we believe that having 
	section 5.2 somehow self-contained could be useful to the readers who are 
	numerically oriented and not interested in the Bethe ansatz details 
	of the result. Therefore, we would prefer, if the referee does not have 
	further objections, to keep section 5.2 as it is.

REFEREE:

	Monte Carlo algorithm: The authors may explain why one cannot simply
	restrict to the sector with N1 = 0 and then re-weight the whole
	sum. Indeed in reference [Phys. Rev. Lett. 110, 257203 (2013)] the
	representative state was chosen to have N1 = 0, as it was for the
	amount of zero momentum strings, therefore the logic should work in
	the same way.

RESPONSE:

	We thank the referee for this useful comment. We agree with the referee 
	that it should be possible to obtain the same results restricting to 
	the sector with N1=0. However, in our work we tried to present a numerical 
	procedure that was as general as possible. In doing so we decided to 
	include in our treatment all the spin sectors, including N1\ne0. The fact 
	that only N1=0 is the only relevant one in the thermodynamic limit 
	is not a necessary ingredient of the algorithm. Moreover, restricting to 
	N1=0 should not provide a significant improvement of the method as 
	the majority of the eigenstates of the XXX chain are in the sector with 
	N1=0;

REFEREE:

	Also would be very nice to have a plot of convergence of the Monte
	Carlo as a function of the Nsteps. One would expect that for larger
	system sizes the number of steps required is smaller if the initial state
	is chosen to be very close to the representative state. How the initial
	state of the algorithm is chosen? Did the authors try to select a proper
	discretization of the thermodynamic saddle point state ? (as done for
	example in [Phys. Rev. A 89, 033605 (2014), J. Phys. A: Math. Theor.
	48, 43FT01 (2015)])

RESPONSE:

	We thank the referee for raising these very important points. First, 
	the chosen initial state is not, in principle, important for the method. 
	The eigenstates sampled at large Monte Carlo times should not be affected 
	by the choice of the initial state. Of course a careful choice of the  
	initial state could could speed up the equilibration of the Monte Carlo. 
	However, we did not perform a sistematic study of that, and for simplicity, 
	we started the Monte Carlo simulations from the ground state of the XXX 
	chain.

	Regarding the possibility of using as initial state the discretization of 
	the saddle point state, this could certainly provide a way to improve 
	the algorithm. However, in our study we preferred not to exploit this 
	possibility in order to provide an unbiased benchmark of our numerical 
	method.

	Finally, regarding the convergence of the Monte Carlo procedure as a 
	function of the number of Monte Carlo steps, the behavior is more 
	complicated than the one described by the referee. First, in our Metropolis 
	step (formula (82) in the manuscript) we chose the proposed new eigenstate 
	|\lambda_j'\rangle completely random. This implies that in general 
	one has

		Re(\epilon'-\epsilon) \sim O(L)

	In turn this implies that if an eigenstate |\lambda_j\rangle corresponds to a 
	local minimum (not the global  one) it can be ''difficult'' for 
	the algorithm to get to a different eigenstate, because the Metropolis 
	transition probability decays exponentially as the system size increases.

	The main consequence is that at fixed number of Monte Carlo steps 
	the error bar increases dramatically upon increasing chain size. This is the 
	reason why we provide only results for L\le 60.

	This problem could be solved by changing the update scheme. The key idea would 
	be to provide in the Metropolis step a proposed new eigenstate |\lambda_j'
	\rangle such that

		\epsilon'-\epsilon \sim O(1)

	This can be clearly achieved if the Bethe number configuration corresponding to 
	|\lambda'_j\rangle are obtained from that of |\lambda_j\rangle by change only 
	``few'' Bethe numbers.

	However, the non triviality of the string content of the generic eigenstates 
	of the XXX chain, makes the implementation of this tricky, and we decided no to 
	do that in the present version of the code.

REFEREE:

	Formula 84 is generically wrong since a generic observable is a more
	involved function of the different string contents of the state (i.e. a
	determinant) and not simply a sum (which is true for the local and
	quasi-local conserved quantities). The authors should then rephrase
	then this point.

RESPONSE:

	We agree with the referee about this point. Below formula 84 we now added 
	the sentence

		"We should stress again that~\eref{ref-form} is true only
		for local or quasi-local observables, while generic observables are more complicated
		functions of the rapidities."

	to stress the importance of locality (and quasi-locality).

REFEREE:

	The authors should compute the expectation values of some simple
	local operators assion for the matrix elements are
	indeed well known (for example in arXiv:1108.6224).

RESPONSE:

	We thank the referee for this very useful comment. In fact, being able to calculate  
	generic correlation functions, besides being interesting on its own, would also provide 
	a more stringent check of the validity of the numerical method. However, 
	the correlation function proposed by the referee corresponds to the local energy 
	that we already provided in the manuscript.

	In order to benchmark our results for higher spins correlation functions we 
	now include in Figure 5 Monte Carlo results for the fluctuations of the local energy 
	density Q_2. Similar to Q4 this  corresponds to a four-spin opearator.

	We could also calculate higher spin correlation functions or some long range 
	correalation functions using known results (similar to 	the one mentioned by the 
	referee). However, this would imply a considerable numerical effort and we believe 
	it goes beyond the scope of the present paper.

REFEREE:

	Once implemented this last point, the ensemble of states created by the
	Monte Carlo algorithm should be enough to reproduce the whole time
	evolution after the quench with the logic of the quench action method
	(see [Phys. Rev. Lett. 110, 257203 (2013), J. Phys. A: Math. Theor.
	48, 43FT01 (2015)]). Can the authors comment on this?

RESPONSE:

	We thank the referee for raising this interesting point. We are in fact aware 
	of the possibility of simulating the full time evolution after a quantum quench 
	starting from the saddle point result. We believe that this would represent 
	a breaktrhough and could provide valuable insights to improve for instance 
	existing numerical methods such as tDMRG. We think that this would deserve a lot 
	of attention on its own and we plan to address it in future work.